---
title: "Chemotherapy Dosing: eGFR Equations vs Measured GFR (OncoGFR 1200)"
author: "Tianqi Ouyang"
date: "2026-05-29"
execute:
  eval: true
  echo: false
  warning: false
  message: false
---

# 1 Setup {.content-hidden}

```{r}
#| label: setup-onco
# Force UTF-8 so unicode (mL/min/1.73m², ±, ≥) renders correctly.
invisible(Sys.setlocale("LC_ALL", "en_US.UTF-8"))

# Lightweight install guard so the project renders on a fresh machine.
.need <- c("readxl","dplyr","tidyr","purrr","tibble","flextable","officer")
.miss <- setdiff(.need, rownames(installed.packages()))
if (length(.miss)) install.packages(.miss, repos = "https://cloud.r-project.org")

library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(flextable)

# Output directory for downloadable Word copies of every table.
docx_dir <- "docs/tables/heatmap"
dir.create(docx_dir, recursive = TRUE, showWarnings = FALSE)
# Absolute path: flextable_to_rmd() can change the working directory mid-render,
# which would break later saves to a relative docx_dir.
docx_dir <- normalizePath(docx_dir, mustWork = TRUE)

save_ft_docx <- function(ft, name) {
  flextable::save_as_docx(ft, path = file.path(docx_dir, paste0(name, ".docx")))
  invisible(ft)
}

# --- Two colouring schemes -------------------------------------------------
# V1: 4-level categorical severity.  <5 yellow · 5–10 orange · 10–20 red · >20 purple.
cat4_fill <- function(x) {
  dplyr::case_when(is.na(x) ~ "#FFFFFF",
                   x <  5  ~ "#FFEDA0",   # yellow
                   x < 10  ~ "#FEB24C",   # orange
                   x <= 20 ~ "#F03B20",   # red
                   TRUE    ~ "#6A51A3")   # purple
}

# V2: continuous gradient across `stops` (low 0 % -> high), saturating at `dom` %.
# (base grDevices, no extra dependency). The exact % is printed in every cell.
ramp_fill <- function(x, stops, dom = 30) {
  f   <- grDevices::colorRamp(stops)
  x2  <- pmin(pmax(x, 0), dom) / dom
  out <- rep("#FFFFFF", length(x))      # NA -> white (no data)
  ok  <- !is.na(x2)
  if (any(ok)) {
    m <- f(x2[ok])
    out[ok] <- grDevices::rgb(m[, 1], m[, 2], m[, 3], maxColorValue = 255)
  }
  out
}
# Yellow -> orange -> red ramp for Part B (low % = yellow, high % = deep red).
CONT_STOPS <- c("#FFEDA0", "#FEB24C", "#F03B20", "#B2182B")
```

# 2 Data Management {.content-hidden}

```{r}
#| label: data-onco
raw <- suppressWarnings(
  readxl::read_excel("Data/OncoGFR1200_MGH_chemo_CJASN.xlsx")
)
dat <- raw %>%
  dplyr::select(-dplyr::any_of(c("...70", "...75"))) %>%
  dplyr::mutate(
    sex     = factor(gender, levels = c(1, 2), labels = c("Male", "Female")),
    bmi_grp = cut(BMI, breaks = c(-Inf, 20, 25, 30, Inf), right = FALSE,
                  labels = c("<20", "20 to <25", "25 to <30", ">=30")),
    # metastasis is complete (0/1) in this dataset — no "Unknown" level.
    met_grp = factor(metastasis, levels = c(0, 1),
                     labels = c("Not metastatic", "Metastatic")),
    age_grp = factor(ifelse(age >= 70, ">=70", "<70"), levels = c("<70", ">=70"))
  )

# --- Equation registry: type + label + index/non-index columns -------------
# Reference = measured GFR. Note: the creatinine CKD-EPI column is the 2021
# (race-free) equation; the hyphen in "EKFCcrcysnon-index" is intentional.
reg <- tibble::tribble(
  ~equation,       ~type,        ~label,             ~index_col,           ~nonindex_col,
  "CKDEPIcr21",    "eGFRcr",     "CKD-EPI 2021",     "CKDEPIcr21index",    "CKDEPIcr21non_index",
  "EKFCcr",        "eGFRcr",     "EKFC 2021",        "EKFCcrindex",        "EKFCcrnon_index",
  "CKDEPIcys",     "eGFRcys",    "CKD-EPI 2012",     "CKDEPIcysindex",     "CKDEPIcysnon_index",
  "EKFCcys",       "eGFRcys",    "EKFC 2023",        "EKFCcysindex",       "EKFCcysnon_index",
  "CKDEPIcrcys21", "eGFRcr-cys", "CKD-EPI 2021",     "CKDEPIcrcys21index", "CKDEPIcrcys21non_index",
  "EKFCcrcys",     "eGFRcr-cys", "EKFC 2023",        "EKFCcrcysindex",     "EKFCcrcysnon-index",
  "CG",            "ClCr",       "Cockcroft-Gault",  "CGindex",            "CGnon_index",
  "CKDEPI4mark",   "Panel",      "CKD-EPI 4-marker", "CKDEPI4mark",        "CKDEPI4marknon_index"
)
MGFR <- c(index = "mGFRindex", nonindex = "mGFRnon_index")

# --- Dosing rules ----------------------------------------------------------
dose_tier <- function(gfr, drug) {           # higher integer = more drug; 0 = avoid
  drug <- tolower(drug)
  if (drug == "cisplatin")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr >= 60 ~ 3, gfr >= 50 ~ 2, gfr >= 40 ~ 1, TRUE ~ 0)
  else if (drug == "methotrexate")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr >= 100 ~ 2, gfr >= 40 ~ 1, TRUE ~ 0)
  else if (drug == "melphalan")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr > 60 ~ 2, gfr >= 30 ~ 1, TRUE ~ 0)
  else stop("unknown drug: ", drug)
}
dose_class_tier <- function(gfr, mgfr, drug) {
  te <- dose_tier(gfr, drug); tr <- dose_tier(mgfr, drug)
  dplyr::case_when(is.na(te) | is.na(tr) ~ NA_character_,
                   te > tr ~ "Overdose", te < tr ~ "Underdose", TRUE ~ "Concordant")
}
# Carboplatin Calvert (AUC 5): over if 5*(eGFR+25) > k_hi*(mGFR+25); under if < k_lo.
# ±20% → k_hi=6, k_lo=4   |   ±15% → k_hi=5.75, k_lo=4.25
carbo_class <- function(gfr, mgfr, k_hi = 6, k_lo = 4) {
  pred <- 5 * (gfr + 25)
  dplyr::case_when(is.na(gfr) | is.na(mgfr) ~ NA_character_,
                   pred > k_hi * (mgfr + 25) ~ "Overdose",
                   pred < k_lo * (mgfr + 25) ~ "Underdose", TRUE ~ "Concordant")
}

TIER_DRUGS  <- c("cisplatin", "methotrexate", "melphalan")
CARBO_BANDS <- list(carboplatin_b20 = c(k_hi = 6, k_lo = 4),
                    carboplatin_b15 = c(k_hi = 5.75, k_lo = 4.25))

# Patient-level classification (carries subgroup columns) for one GFR scale.
build_class_long <- function(scale = c("index", "nonindex")) {
  scale  <- match.arg(scale)
  col_of <- function(eq) {
    r <- reg[reg$equation == eq, ]
    if (scale == "index") r$index_col[[1]] else r$nonindex_col[[1]]
  }
  mgfr <- dat[[MGFR[[scale]]]]
  base <- dplyr::select(dat, sex, bmi_grp, met_grp, age_grp)
  tier <- purrr::map_dfr(reg$equation, function(eq) {
    g <- dat[[col_of(eq)]]
    purrr::map_dfr(TIER_DRUGS, function(dr)
      dplyr::bind_cols(base, tibble::tibble(equation = eq, drug = dr,
                                            cls = dose_class_tier(g, mgfr, dr))))
  })
  carbo <- purrr::map_dfr(reg$equation, function(eq) {
    g <- dat[[col_of(eq)]]
    purrr::map_dfr(names(CARBO_BANDS), function(bn) {
      k <- CARBO_BANDS[[bn]]
      dplyr::bind_cols(base, tibble::tibble(equation = eq, drug = bn,
                                            cls = carbo_class(g, mgfr, k[["k_hi"]], k[["k_lo"]])))
    })
  })
  dplyr::bind_rows(tier, carbo)
}
cl_index    <- build_class_long("index")
cl_nonindex <- build_class_long("nonindex")

# Internal (syntactic) column keys + their display labels.
SUB_COLS <- c("Overall", "Female", "Male", "bmi1", "bmi2", "bmi3", "bmi4",
              "met_no", "met_yes", "age_lt70", "age_ge70")
SUB_LABS <- c(Overall = "Overall", Female = "Female", Male = "Male",
              bmi1 = "<20", bmi2 = "20 to <25", bmi3 = "25 to <30", bmi4 = ">=30",
              met_no = "Not metastatic", met_yes = "Metastatic",
              age_lt70 = "<70", age_ge70 = "≥70")

# Build the equation × subgroup % table for one drug and direction.
subgroup_table <- function(cl, drug_id, direction) {
  s <- cl %>% dplyr::filter(drug == drug_id, !is.na(cls))
  pct <- function(d) if (nrow(d)) round(mean(d$cls == direction) * 100, 1) else NA_real_
  lvl <- function(d, col, value) pct(d[!is.na(d[[col]]) & d[[col]] == value, ])
  per_eq <- function(eq) {
    e <- s %>% dplyr::filter(equation == eq)
    tibble::tibble(
      equation         = eq,
      Overall          = pct(e),
      Female           = lvl(e, "sex", "Female"),
      Male             = lvl(e, "sex", "Male"),
      bmi1    = lvl(e, "bmi_grp", "<20"),
      bmi2    = lvl(e, "bmi_grp", "20 to <25"),
      bmi3    = lvl(e, "bmi_grp", "25 to <30"),
      bmi4    = lvl(e, "bmi_grp", ">=30"),
      met_no   = lvl(e, "met_grp", "Not metastatic"),
      met_yes  = lvl(e, "met_grp", "Metastatic"),
      age_lt70 = lvl(e, "age_grp", "<70"),
      age_ge70 = lvl(e, "age_grp", ">=70")
    )
  }
  purrr::map_dfr(reg$equation, per_eq) %>%
    dplyr::left_join(dplyr::select(reg, equation, type, label), by = "equation") %>%
    dplyr::mutate(type = factor(type, levels = unique(reg$type))) %>%
    dplyr::arrange(type, match(equation, reg$equation)) %>%
    dplyr::select(type, label, dplyr::all_of(SUB_COLS))
}

# Render one heatmap flextable. `fill_fn(v)` -> per-cell hex; `dark_fn(v)` -> TRUE
# where the fill is dark enough to need white text.
make_tl_ft <- function(tab, title, fill_fn, dark_fn) {
  ncol_total <- length(c("type", "label", SUB_COLS))   # 11
  ft <- flextable::flextable(tab, col_keys = c("type", "label", SUB_COLS))
  ft <- flextable::set_header_labels(ft,
    values = c(list(type = "eGFR type", label = "Equation"), as.list(SUB_LABS)))
  # Group-spanner header row (Sex / BMI / Metastatic disease).
  ft <- flextable::add_header_row(ft, top = TRUE,
    values    = c("", "", "", "Sex", "BMI category (kg/m²)", "Metastatic disease", "Age (years)"),
    colwidths = c(1, 1, 1, 2, 4, 2, 2))
  # Title row across the whole table ("OVERDOSE (%)" / "UNDERDOSE (%)").
  ft <- flextable::add_header_row(ft, top = TRUE, values = title, colwidths = ncol_total)
  for (cc in SUB_COLS) {
    v  <- tab[[cc]]
    ft <- flextable::bg(ft, j = cc, bg = fill_fn(v), part = "body")
    # white text on dark cells — scalar colour on specific rows (avoids a
    # vectorised colour arg, which is fragile under R 4.4).
    dark <- which(dark_fn(v))
    if (length(dark)) ft <- flextable::color(ft, i = dark, j = cc, color = "white", part = "body")
  }
  ft <- flextable::merge_v(ft, j = "type", part = "body")
  ft <- flextable::valign(ft, j = "type", valign = "top", part = "body")
  ft <- flextable::align(ft, j = SUB_COLS, align = "center", part = "all")
  ft <- flextable::align(ft, i = 2, align = "center", part = "header")  # spanner row
  ft <- flextable::align(ft, i = 1, align = "left",   part = "header")  # title row
  ft <- flextable::bold(ft, part = "header")
  ft <- flextable::fontsize(ft, size = 9, part = "all")
  ft <- flextable::fontsize(ft, i = 1, size = 11, part = "header")
  flextable::set_table_properties(ft, layout = "autofit", width = 1)  # renderer-side sizing
}

# Build the flextable for one (cl, drug, direction) under colour scheme `version`.
build_ft <- function(cl, drug_id, direction, version) {
  tab   <- subgroup_table(cl, drug_id, direction)
  title <- paste0(toupper(direction), " (%)")
  if (version == "cat") {
    fill_fn <- cat4_fill
    dark_fn <- function(x) !is.na(x) & x >= 10            # red & purple -> white text
  } else {
    fill_fn <- function(x) ramp_fill(x, CONT_STOPS, 30)   # yellow->red, saturate at 30 %
    dark_fn <- function(x) !is.na(x) & x >= 20            # white text on the red end
  }
  make_tl_ft(tab, title, fill_fn, dark_fn)
}

DRUGS <- list(
  list(id = "carboplatin_b20", lab = "Carboplatin (±20%)"),
  list(id = "carboplatin_b15", lab = "Carboplatin (±15%)"),
  list(id = "cisplatin",       lab = "Cisplatin"),
  list(id = "methotrexate",    lab = "Methotrexate"),
  list(id = "melphalan",       lab = "Melphalan")
)

# Emit every drug × direction for one scale/version (call inside a results='asis'
# chunk). Saves each table to Word and prints it with a download link.
emit_scale <- function(cl, version, scale, scale_label) {
  cat("\n\n### ", scale_label, "\n\n", sep = "")
  for (d in DRUGS) {
    cat("\n\n#### ", d$lab, "\n\n", sep = "")
    for (dir in c("Overdose", "Underdose")) {
      fname <- paste0(version, "_", scale, "_", d$id, "_", tolower(dir))
      ft <- build_ft(cl, d$id, dir, version)
      save_ft_docx(ft, fname)
      cat(flextable::flextable_to_rmd(ft))
      cat("\n\n")
    }
    cat(sprintf(
      "[📄 Overdose Word](../tables/heatmap/%1$s_%2$s_%3$s_overdose.docx) · [📄 Underdose Word](../tables/heatmap/%1$s_%2$s_%3$s_underdose.docx)\n\n",
      version, scale, d$id))
  }
}
```

# Overview

For every patient we compute the chemotherapy dose each **estimated-GFR (eGFR) equation** would
prescribe and compare it to the dose that **measured GFR (mGFR, ⁵¹Cr-EDTA clearance)** — the gold
standard — would prescribe. Each cell is the **percent of patients mis-dosed** in that subgroup:

- **OVERDOSE (%)** — the equation would give *more* drug than mGFR warrants.
- **UNDERDOSE (%)** — the equation would give *less* drug than mGFR warrants.

**Cell colour** — the number in each cell is the **share of patients in that subgroup whose dose
would be wrong in that direction** (over- or under-dosed vs mGFR). The same tables are shown in
**two colour versions** so you can compare:

- **Part A — categorical (4 severity bands):**
  <span style="background:#FFEDA0">&nbsp;yellow &lt; 5 %&nbsp;</span> ·
  <span style="background:#FEB24C">&nbsp;orange 5–10 %&nbsp;</span> ·
  <span style="background:#F03B20;color:white">&nbsp;red 10–20 %&nbsp;</span> ·
  <span style="background:#6A51A3;color:white">&nbsp;purple &gt; 20 %&nbsp;</span>.
- **Part B — continuous:** a shade of the value,
  <span style="background:#FFEDA0">&nbsp;yellow = 0 %&nbsp;</span> →
  <span style="background:#FEB24C">&nbsp;orange&nbsp;</span> →
  <span style="background:#B2182B;color:white">&nbsp;deep red&nbsp;</span>, saturating (worst) at
  **≥ 30 %**.

Both are severity scales (higher % = worse) applied to over- and under-dose alike; the exact
percentage is printed in every cell.

Columns give the **Overall** rate and rates within **Sex**, **BMI category**, **Metastatic
disease**, and **Age** (<70 vs ≥70) subgroups. Results are shown on both the **indexed** (mL/min/1.73 m²) and **non-indexed**
(absolute mL/min) GFR scale; mGFR is always taken on the same scale as the equation. All 1,200
patients are used for every drug (a counterfactual — the file has no administered doses).

**Dosing rules** — one row per GFR band. *Tier* is the ordered dose code used internally
(higher = more drug; 0 = avoid / forced reduce).

| Drug | GFR (mL/min) | Dose | Tier |
|---|---|---|:--:|
| **Cisplatin** | ≥ 60 | Full dose | 3 |
|  | 50 – 59 | −25 % | 2 |
|  | 40 – 49 | −50 % | 1 |
|  | < 40 | Avoid | 0 |
| **Methotrexate** (CNS, 8 g/m²) | ≥ 100 | Full dose | 2 |
|  | 40 – 99 | Reduced (proportional) | 1 |
|  | < 40 | Avoid | 0 |
| **Melphalan** (200 mg/m²) | > 60 | Full, 200 mg/m² | 2 |
|  | 30 – 59 | 140 mg/m² | 1 |
|  | < 30 | Forced reduce, 140 mg/m² | 0 |
| **Carboplatin** (Calvert) | continuous (AUC) | dose = target AUC × (GFR + 25), target AUC 5; over-/under-dose set by a ±20 % or ±15 % band — **see note below** | — |

::: callout-note
## How carboplatin over- and under-dosing are defined

Carboplatin is **not** dosed in GFR tiers like the other three drugs — it is dosed to a target drug
exposure with the **Calvert formula**: **dose (mg) = target AUC × (GFR + 25)**, with target
**AUC = 5**. Because the dose is *directly proportional* to the GFR entered, plugging an eGFR
equation into the formula instead of the true (measured) GFR changes the dose by the **same
percentage** that the equation mis-estimates GFR. We flag only a *clinically meaningful* error,
using a tolerance band around the AUC-5 target:

- **Overdose** — the eGFR-based dose is larger than the dose from **measured GFR at AUC 6**. Since
  AUC 6 is 20 % above AUC 5, the equation would deliver **> 20 % too much** carboplatin.
- **Underdose** — the eGFR-based dose is smaller than the dose from **measured GFR at AUC 4** (20 %
  below AUC 5) → **> 20 % too little**.

That ±20 % band is reported as **`carboplatin_b20`**; **`carboplatin_b15`** repeats the same test
with a tighter **±15 %** band (AUC 5.75 and 4.25). Both are shown because the analysis plan asked
for 15 % and 20 % thresholds. (These definitions come from the eligibility-cutoffs document.)
:::

::: callout-note
"Metastatic disease — Unknown" from the source template is omitted because metastatic status is
complete (no missing) in this dataset. The creatinine equation shown is **CKD-EPI 2021** (the
race-free version present in the data).
:::

# Variable codebook

Every field used to build the tables. **Inputs** are columns read from
`OncoGFR1200_MGH_chemo_CJASN.xlsx` (N = 1,200); **derived** fields are computed in this analysis.

### Kidney-function inputs (eGFR / mGFR)

Each equation is stored twice — an **index** column (mL/min/1.73 m², BSA-normalised) and a
**non-index** column (absolute mL/min ≈ indexed × BSA ⁄ 1.73). Measured GFR is the reference all
equations are compared against; comparisons are made on the matching scale.

| eGFR type | Equation (label) | Index column | Non-index column |
|---|---|---|---|
| reference | Measured GFR (⁵¹Cr-EDTA) | `mGFRindex` | `mGFRnon_index` |
| eGFRcr | CKD-EPI 2021 (creatinine) | `CKDEPIcr21index` | `CKDEPIcr21non_index` |
| eGFRcr | EKFC 2021 (creatinine) | `EKFCcrindex` | `EKFCcrnon_index` |
| eGFRcys | CKD-EPI 2012 (cystatin C) | `CKDEPIcysindex` | `CKDEPIcysnon_index` |
| eGFRcys | EKFC 2023 (cystatin C) | `EKFCcysindex` | `EKFCcysnon_index` |
| eGFRcr-cys | CKD-EPI 2021 (Cr–Cys) | `CKDEPIcrcys21index` | `CKDEPIcrcys21non_index` |
| eGFRcr-cys | EKFC 2023 (Cr–Cys) | `EKFCcrcysindex` | `EKFCcrcysnon-index` |
| ClCr | Cockcroft–Gault | `CGindex` | `CGnon_index` |
| Panel | CKD-EPI 4-marker (Cr, Cys, β2M, BTP) | `CKDEPI4mark` | `CKDEPI4marknon_index` |

*Underlying labs (`Cr`, `CysC`, `B2M`, `BTP`) and anthropometrics (`weight`, `height`, `BSA`,
`age`, `gender`) were used upstream to compute these equation values; this analysis reads only the
equation columns plus `gender`, `BMI`, and `metastasis` (below).*

### Subgroup variables (table columns)

| Derived field | Source column | Levels / definition |
|---|---|---|
| `Overall` | — | all 1,200 patients |
| `sex` | `gender` (1, 2) | Male (1); Female (2) |
| `bmi_grp` | `BMI` (kg/m²) | `<20`; `20 to <25`; `25 to <30`; `≥30` |
| `met_grp` | `metastasis` (0, 1) | Not metastatic (0); Metastatic (1) |
| `age_grp` | `age` (years) | `<70`; `≥70` (binary cut-off at 70) |

### Computed classification fields

| Field | Values | Meaning |
|---|---|---|
| `drug` | `cisplatin`, `methotrexate`, `melphalan`, `carboplatin_b20`, `carboplatin_b15` | which drug's dosing rule was applied. `carboplatin_b20` and `carboplatin_b15` are the **same** carboplatin analysis run with a ±20 % vs ±15 % tolerance band (see the carboplatin note under *Overview*). |
| `dose_tier` | 0–3 | ordered dose level from the rule (see *Dosing rules*); higher = more drug, 0 = avoid |
| `dose_class` (`cls`) | Overdose / Underdose / Concordant | equation's tier vs mGFR's tier (Calvert band for carboplatin) |
| `pct_over` | 0–100 | % of the subgroup classified Overdose → **OVERDOSE (%)** cells |
| `pct_under` | 0–100 | % of the subgroup classified Underdose → **UNDERDOSE (%)** cells |
| `equation` / `type` / `label` | — | equation id / eGFR family / display name (see table above) |

Cell colour: two versions of the same tables — **Part A** categorical (yellow <5 / orange 5–10 / red 10–20 / purple >20) and **Part B** continuous yellow→red (worst at ≥30 %). Higher % = worse in both; the number is printed in every cell.

# Part A — Categorical colour (4 levels)

Severity bands (same scale for over- and under-dose):
<span style="background:#FFEDA0">&nbsp;yellow &lt; 5 %&nbsp;</span> ·
<span style="background:#FEB24C">&nbsp;orange 5–10 %&nbsp;</span> ·
<span style="background:#F03B20;color:white">&nbsp;red 10–20 %&nbsp;</span> ·
<span style="background:#6A51A3;color:white">&nbsp;purple &gt; 20 %&nbsp;</span>

```{r}
#| label: partA-index
#| results: asis
emit_scale(cl_index, "cat", "index", "Indexed GFR (mL/min/1.73 m²)")
```

```{r}
#| label: partA-nonindex
#| results: asis
emit_scale(cl_nonindex, "cat", "nonindex", "Non-indexed GFR (absolute mL/min)")
```

# Part B — Continuous colour (yellow → red)

Shade ∝ value, saturating at 30 %:
<span style="background:#FFEDA0">&nbsp;yellow = 0 %&nbsp;</span> →
<span style="background:#FEB24C">&nbsp;orange&nbsp;</span> →
<span style="background:#B2182B;color:white">&nbsp;deep red ≥ 30 %&nbsp;</span> (darker = worse). The
exact % is printed in every cell.

```{r}
#| label: partB-index
#| results: asis
emit_scale(cl_index, "cont", "index", "Indexed GFR (mL/min/1.73 m²)")
```

```{r}
#| label: partB-nonindex
#| results: asis
emit_scale(cl_nonindex, "cont", "nonindex", "Non-indexed GFR (absolute mL/min)")
```


# 5 Notes

- **Reference.** Measured GFR (⁵¹Cr-EDTA) is the gold standard; each equation is compared to mGFR on
  the matching scale (indexed vs indexed, non-indexed vs non-indexed).
- **Scale.** Calvert carboplatin dosing was derived on **absolute (non-indexed)** GFR (Section 4 is
  the clinically conventional view for carboplatin); indexed GFR (Section 3) is the conventional
  scale for CKD-staging cutoffs.
- **Direction.** Creatinine-based equations tend to **overdose** (they overestimate GFR, especially
  at low BMI / low muscle mass); cystatin-based equations tend to **underdose**.
- **Counterfactual.** No administered doses exist in this file, so every drug rule is applied to all
  1,200 patients to estimate misclassification rates, not actual prescribing errors.
