---
title: "Chemotherapy Dosing: eGFR Equations vs Measured GFR (OncoGFR 1200)"
author: "Tianqi Ouyang"
date: "2026-05-29"
execute:
  eval: true
  echo: false
  warning: false
  message: false
format:
  html:
    fig-dpi: 96
    fig-format: png
    code-fold: true
---

# 1 Setup {.content-hidden}

```{r}
#| label: setup-onco
# Force UTF-8 so unicode (mL/min/1.73m², ±, ≥) renders correctly.
invisible(Sys.setlocale("LC_ALL", "en_US.UTF-8"))

# Lightweight install guard so the project renders on a fresh machine.
.need <- c("readxl","dplyr","tidyr","purrr","tibble","flextable","officer","ggplot2")
.miss <- setdiff(.need, rownames(installed.packages()))
if (length(.miss)) install.packages(.miss, repos = "https://cloud.r-project.org")

library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(flextable)
library(ggplot2)

# Output directory for downloadable Word copies of every table.
docx_dir <- "docs/tables/heatmap"
dir.create(docx_dir, recursive = TRUE, showWarnings = FALSE)
# Absolute path: flextable_to_rmd() can change the working directory mid-render,
# which would break later saves to a relative docx_dir.
docx_dir <- normalizePath(docx_dir, mustWork = TRUE)

save_ft_docx <- function(ft, name) {
  flextable::save_as_docx(ft, path = file.path(docx_dir, paste0(name, ".docx")))
  invisible(ft)
}

# --- Two colouring schemes -------------------------------------------------
# V1: 4-level categorical severity.  <5 yellow · 5–10 orange · 10–20 red · >20 purple.
cat4_fill <- function(x) {
  dplyr::case_when(is.na(x) ~ "#FFFFFF",
                   x <  5  ~ "#FFEDA0",   # yellow
                   x < 10  ~ "#FEB24C",   # orange
                   x <= 20 ~ "#F03B20",   # red
                   TRUE    ~ "#6A51A3")   # purple
}

# V2: continuous gradient across `stops` (low 0 % -> high), saturating at `dom` %.
# (base grDevices, no extra dependency). The exact % is printed in every cell.
ramp_fill <- function(x, stops, dom = 30) {
  f   <- grDevices::colorRamp(stops)
  x2  <- pmin(pmax(x, 0), dom) / dom
  out <- rep("#FFFFFF", length(x))      # NA -> white (no data)
  ok  <- !is.na(x2)
  if (any(ok)) {
    m <- f(x2[ok])
    out[ok] <- grDevices::rgb(m[, 1], m[, 2], m[, 3], maxColorValue = 255)
  }
  out
}
# Yellow -> orange -> red ramp for Part B (low % = yellow, high % = deep red).
CONT_STOPS <- c("#FFEDA0", "#FEB24C", "#F03B20", "#B2182B")
```

# 2 Data Management {.content-hidden}

```{r}
#| label: data-onco
raw <- suppressWarnings(
  readxl::read_excel("Data/OncoGFR1200_MGH_chemo_CJASN.xlsx")
)
dat <- raw %>%
  dplyr::select(-dplyr::any_of(c("...70", "...75"))) %>%
  dplyr::mutate(
    sex     = factor(gender, levels = c(1, 2), labels = c("Male", "Female")),
    bmi_grp = cut(BMI, breaks = c(-Inf, 20, 25, 30, Inf), right = FALSE,
                  labels = c("<20", "20 to <25", "25 to <30", ">=30")),
    # metastasis is complete (0/1) in this dataset — no "Unknown" level.
    met_grp = factor(metastasis, levels = c(0, 1),
                     labels = c("Not metastatic", "Metastatic")),
    age_grp = factor(ifelse(age >= 70, ">=70", "<70"), levels = c("<70", ">=70")),
    # ECOG performance status; only 8 patients are ECOG 3, so lump 2+3 -> "≥2".
    ecog_grp = factor(ifelse(ecog >= 2, "≥2", as.character(ecog)), levels = c("0", "1", "≥2")),
    # LOESS x-axes: eGFRcys (CKD-EPI 2012) vs eGFRcr (CKD-EPI 2021). The RATIO is
    # scale-invariant; the DIFFERENCE is not, so it is taken on the indexed scale
    # (mL/min/1.73 m²). Both use the index columns.
    egfr_ratio = CKDEPIcysindex / CKDEPIcr21index,
    egfr_diff  = CKDEPIcysindex - CKDEPIcr21index,
    # % difference of the two eGFRs relative to creatinine; scale-invariant.
    egfr_pctdiff = (CKDEPIcr21index - CKDEPIcysindex) / CKDEPIcr21index * 100
  )

# Body-composition subset (465 patients, matched on `protocol`). Adds a Prado
# sarcopenia group and an in_bodycomp flag; NA for the other 735 patients.
.bc <- suppressWarnings(readxl::read_excel("Data/OncoGFR_MGH_final_0428.xlsx"))
dat <- dat %>% dplyr::mutate(
  in_bodycomp = protocol %in% .bc$protocol,
  sarc_grp = factor(.bc$sarcopenia_prado[match(protocol, .bc$protocol)],
                    levels = c(0, 1), labels = c("Non-sarcopenic", "Sarcopenic")),
  muscle_z = .bc$z_score[match(protocol, .bc$protocol)]   # muscle (SMI) Z-score
)

# --- Equation registry: type + label + index/non-index columns -------------
# Reference = measured GFR. Note: the creatinine CKD-EPI column is the 2021
# (race-free) equation; the hyphen in "EKFCcrcysnon-index" is intentional.
reg <- tibble::tribble(
  ~equation,       ~type,        ~label,             ~index_col,           ~nonindex_col,
  "CKDEPIcr21",    "eGFRcr",     "CKD-EPI 2021",     "CKDEPIcr21index",    "CKDEPIcr21non_index",
  "EKFCcr",        "eGFRcr",     "EKFC 2021",        "EKFCcrindex",        "EKFCcrnon_index",
  "CKDEPIcys",     "eGFRcys",    "CKD-EPI 2012",     "CKDEPIcysindex",     "CKDEPIcysnon_index",
  "EKFCcys",       "eGFRcys",    "EKFC 2023",        "EKFCcysindex",       "EKFCcysnon_index",
  "CKDEPIcrcys21", "eGFRcr-cys", "CKD-EPI 2021",     "CKDEPIcrcys21index", "CKDEPIcrcys21non_index",
  "EKFCcrcys",     "eGFRcr-cys", "EKFC 2023",        "EKFCcrcysindex",     "EKFCcrcysnon-index",
  "CG",            "ClCr",       "Cockcroft-Gault",  "CGindex",            "CGnon_index",
  "CKDEPI4mark",   "Panel",      "CKD-EPI 4-marker", "CKDEPI4mark",        "CKDEPI4marknon_index"
)
MGFR <- c(index = "mGFRindex", nonindex = "mGFRnon_index")

# --- Dosing rules ----------------------------------------------------------
dose_tier <- function(gfr, drug) {           # higher integer = more drug; 0 = avoid
  drug <- tolower(drug)
  if (drug == "cisplatin")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr >= 60 ~ 3, gfr >= 50 ~ 2, gfr >= 40 ~ 1, TRUE ~ 0)
  else if (drug == "methotrexate")   # ADDIKD: ≥60 full(3); 45-<60 −25%(2); 30-<45 −50%(1); <30 avoid(0)
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr >= 60 ~ 3, gfr >= 45 ~ 2, gfr >= 30 ~ 1, TRUE ~ 0)
  else if (drug == "melphalan")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr > 60 ~ 2, gfr >= 30 ~ 1, TRUE ~ 0)
  else stop("unknown drug: ", drug)
}
dose_class_tier <- function(gfr, mgfr, drug) {
  te <- dose_tier(gfr, drug); tr <- dose_tier(mgfr, drug)
  dplyr::case_when(is.na(te) | is.na(tr) ~ NA_character_,
                   te > tr ~ "Overdose", te < tr ~ "Underdose", TRUE ~ "Concordant")
}
# Carboplatin Calvert (AUC 5): over if 5*(eGFR+25) > k_hi*(mGFR+25); under if < k_lo.
# ±20% → k_hi=6, k_lo=4   |   ±15% → k_hi=5.75, k_lo=4.25
carbo_class <- function(gfr, mgfr, k_hi = 6, k_lo = 4) {
  pred <- 5 * (gfr + 25)
  dplyr::case_when(is.na(gfr) | is.na(mgfr) ~ NA_character_,
                   pred > k_hi * (mgfr + 25) ~ "Overdose",
                   pred < k_lo * (mgfr + 25) ~ "Underdose", TRUE ~ "Concordant")
}

TIER_DRUGS  <- c("cisplatin", "methotrexate", "melphalan")
CARBO_BANDS <- list(carboplatin_b20 = c(k_hi = 6, k_lo = 4),
                    carboplatin_b15 = c(k_hi = 5.75, k_lo = 4.25))

# Patient-level classification (carries subgroup columns) for one GFR scale.
# `data` lets a subset (e.g. the 465 body-comp cohort) reuse the same logic.
build_class_long <- function(scale = c("index", "nonindex"), data = dat) {
  scale  <- match.arg(scale)
  col_of <- function(eq) {
    r <- reg[reg$equation == eq, ]
    if (scale == "index") r$index_col[[1]] else r$nonindex_col[[1]]
  }
  mgfr <- data[[MGFR[[scale]]]]
  base <- dplyr::select(data, sex, bmi_grp, met_grp, age_grp, ecog_grp,
                        egfr_ratio, egfr_diff, egfr_pctdiff, sarc_grp, muscle_z)
  tier <- purrr::map_dfr(reg$equation, function(eq) {
    g <- data[[col_of(eq)]]
    purrr::map_dfr(TIER_DRUGS, function(dr)
      dplyr::bind_cols(base, tibble::tibble(equation = eq, drug = dr,
                                            cls = dose_class_tier(g, mgfr, dr))))
  })
  carbo <- purrr::map_dfr(reg$equation, function(eq) {
    g <- data[[col_of(eq)]]
    purrr::map_dfr(names(CARBO_BANDS), function(bn) {
      k <- CARBO_BANDS[[bn]]
      dplyr::bind_cols(base, tibble::tibble(equation = eq, drug = bn,
                                            cls = carbo_class(g, mgfr, k[["k_hi"]], k[["k_lo"]])))
    })
  })
  dplyr::bind_rows(tier, carbo)
}
cl_index    <- build_class_long("index")
cl_nonindex <- build_class_long("nonindex")

# 465-patient body-composition subset (adds the Sarcopenia column group).
dat_465         <- dat %>% dplyr::filter(in_bodycomp)
cl_index_465    <- build_class_long("index",    dat_465)
cl_nonindex_465 <- build_class_long("nonindex", dat_465)

# Internal (syntactic) column keys + their display labels.
SUB_COLS <- c("Overall", "Female", "Male", "bmi1", "bmi2", "bmi3", "bmi4",
              "met_no", "met_yes", "age_lt70", "age_ge70")
SUB_LABS <- c(Overall = "Overall", Female = "Female", Male = "Male",
              bmi1 = "<20", bmi2 = "20 to <25", bmi3 = "25 to <30", bmi4 = ">=30",
              met_no = "Not metastatic", met_yes = "Metastatic",
              age_lt70 = "<70", age_ge70 = "≥70")

# Extended set for the 465 body-comp subset: adds a Sarcopenia (Prado) group.
SUB_COLS_465 <- c(SUB_COLS, "sarc_no", "sarc_yes")
SUB_LABS_465 <- c(SUB_LABS, sarc_no = "Non-sarcopenic", sarc_yes = "Sarcopenic")

# Build the equation × subgroup % table for one drug and direction.
# extra = TRUE appends the Sarcopenia (Prado) columns (465 body-comp subset).
subgroup_table <- function(cl, drug_id, direction, extra = FALSE) {
  s <- cl %>% dplyr::filter(drug == drug_id, !is.na(cls))
  pct <- function(d) if (nrow(d)) round(mean(d$cls == direction) * 100, 1) else NA_real_
  lvl <- function(d, col, value) pct(d[!is.na(d[[col]]) & d[[col]] == value, ])
  per_eq <- function(eq) {
    e <- s %>% dplyr::filter(equation == eq)
    out <- tibble::tibble(
      equation         = eq,
      Overall          = pct(e),
      Female           = lvl(e, "sex", "Female"),
      Male             = lvl(e, "sex", "Male"),
      bmi1    = lvl(e, "bmi_grp", "<20"),
      bmi2    = lvl(e, "bmi_grp", "20 to <25"),
      bmi3    = lvl(e, "bmi_grp", "25 to <30"),
      bmi4    = lvl(e, "bmi_grp", ">=30"),
      met_no   = lvl(e, "met_grp", "Not metastatic"),
      met_yes  = lvl(e, "met_grp", "Metastatic"),
      age_lt70 = lvl(e, "age_grp", "<70"),
      age_ge70 = lvl(e, "age_grp", ">=70")
    )
    if (extra) {
      out$sarc_no  <- lvl(e, "sarc_grp", "Non-sarcopenic")
      out$sarc_yes <- lvl(e, "sarc_grp", "Sarcopenic")
    }
    out
  }
  cols <- if (extra) SUB_COLS_465 else SUB_COLS
  purrr::map_dfr(reg$equation, per_eq) %>%
    dplyr::left_join(dplyr::select(reg, equation, type, label), by = "equation") %>%
    dplyr::mutate(type = factor(type, levels = unique(reg$type))) %>%
    dplyr::arrange(type, match(equation, reg$equation)) %>%
    dplyr::select(type, label, dplyr::all_of(cols))
}

# Render one heatmap flextable. `fill_fn(v)` -> per-cell hex; `dark_fn(v)` -> TRUE
# where the fill is dark enough to need white text.
make_tl_ft <- function(tab, title, fill_fn, dark_fn, extra = FALSE) {
  cols  <- if (extra) SUB_COLS_465 else SUB_COLS
  labs  <- if (extra) SUB_LABS_465 else SUB_LABS
  spanv <- c("", "", "", "Sex", "BMI category (kg/m²)", "Metastatic disease", "Age (years)")
  spanw <- c(1, 1, 1, 2, 4, 2, 2)
  if (extra) { spanv <- c(spanv, "Sarcopenia (Prado)"); spanw <- c(spanw, 2) }
  ncol_total <- length(c("type", "label", cols))
  ft <- flextable::flextable(tab, col_keys = c("type", "label", cols))
  ft <- flextable::set_header_labels(ft,
    values = c(list(type = "eGFR type", label = "Equation"), as.list(labs)))
  # Group-spanner header row (Sex / BMI / Metastatic / Age [/ Sarcopenia]).
  ft <- flextable::add_header_row(ft, top = TRUE, values = spanv, colwidths = spanw)
  # Title row across the whole table ("OVERDOSE (%)" / "UNDERDOSE (%)").
  ft <- flextable::add_header_row(ft, top = TRUE, values = title, colwidths = ncol_total)
  for (cc in cols) {
    v  <- tab[[cc]]
    ft <- flextable::bg(ft, j = cc, bg = fill_fn(v), part = "body")
    # white text on dark cells — scalar colour on specific rows (avoids a
    # vectorised colour arg, which is fragile under R 4.4).
    dark <- which(dark_fn(v))
    if (length(dark)) ft <- flextable::color(ft, i = dark, j = cc, color = "white", part = "body")
  }
  ft <- flextable::merge_v(ft, j = "type", part = "body")
  ft <- flextable::valign(ft, j = "type", valign = "top", part = "body")
  ft <- flextable::align(ft, j = cols, align = "center", part = "all")
  ft <- flextable::align(ft, i = 2, align = "center", part = "header")  # spanner row
  ft <- flextable::align(ft, i = 1, align = "left",   part = "header")  # title row
  ft <- flextable::bold(ft, part = "header")
  ft <- flextable::fontsize(ft, size = 9, part = "all")
  ft <- flextable::fontsize(ft, i = 1, size = 11, part = "header")
  flextable::set_table_properties(ft, layout = "autofit", width = 1)  # renderer-side sizing
}

# Build the flextable for one (cl, drug, direction) under colour scheme `version`.
build_ft <- function(cl, drug_id, direction, version, extra = FALSE) {
  tab   <- subgroup_table(cl, drug_id, direction, extra)
  title <- paste0(toupper(direction), " (%)")
  if (version == "cat") {
    fill_fn <- cat4_fill
    dark_fn <- function(x) !is.na(x) & x >= 10            # red & purple -> white text
  } else {
    fill_fn <- function(x) ramp_fill(x, CONT_STOPS, 30)   # yellow->red, saturate at 30 %
    dark_fn <- function(x) !is.na(x) & x >= 20            # white text on the red end
  }
  make_tl_ft(tab, title, fill_fn, dark_fn, extra)
}

DRUGS <- list(
  list(id = "carboplatin_b20", lab = "Carboplatin (±20%)"),
  list(id = "carboplatin_b15", lab = "Carboplatin (±15%)"),
  list(id = "cisplatin",       lab = "Cisplatin"),
  list(id = "methotrexate",    lab = "Methotrexate"),
  list(id = "melphalan",       lab = "Melphalan")
)
DRUG_IDS  <- vapply(DRUGS, `[[`, "", "id")
DRUG_LABS <- vapply(DRUGS, `[[`, "", "lab")

# Equations shown in the LOESS panels — all 8. NB: the CKD-EPI Cys 2012 and
# CKD-EPI Cr 2021 panels are partly self-referential (they define the x-axis
# ratio/difference), so interpret those two with care.
LOESS_EQS <- reg$equation

# LOESS: one plot per equation; the 5 drugs are overlaid as colour-coded LOESS
# smooths of P(over/under-dose) vs an x variable (binary 0/1 -> % via mean).
# `xvar`/`xlab`/`xlim` default to the eGFR ratio; the difference set overrides them.
loess_plot <- function(cl, eq, direction, xvar = "egfr_ratio",
                       xlab = "eGFR ratio  (CKD-EPI Cys 2012 / Cr 2021)",
                       xlim = c(0.4, 1.4)) {
  r   <- reg[reg$equation == eq, ]
  ttl <- paste0(r$type, " · ", r$label)
  d <- cl %>%
    dplyr::filter(equation == eq, !is.na(cls), !is.na(.data[[xvar]])) %>%
    dplyr::mutate(y = as.integer(cls == direction) * 100,
                  Drug = factor(drug, levels = DRUG_IDS, labels = DRUG_LABS))
  ggplot2::ggplot(d, ggplot2::aes(.data[[xvar]], y, colour = Drug, fill = Drug)) +
    ggplot2::geom_smooth(method = "loess", formula = y ~ x, span = 0.9,
                         se = TRUE, alpha = 0.10, linewidth = 0.9) +
    ggplot2::coord_cartesian(xlim = xlim, ylim = c(0, 100)) +
    ggplot2::labs(x = xlab, y = paste0("% ", tolower(direction), "d"),
                  title = ttl, colour = NULL, fill = NULL) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "right",
                   plot.title = ggplot2::element_text(face = "bold", size = 11))
}
# X-axis for the difference LOESS set (indexed Cys − Cr, mL/min/1.73 m²).
DIFF_XLAB <- "eGFR difference  (CKD-EPI Cys 2012 − Cr 2021, mL/min/1.73 m²)"
DIFF_XLIM <- c(-50, 20)
# X-axis for the % difference LOESS set ((Cr − Cys)/Cr × 100; scale-invariant).
PCTD_XLAB <- "eGFR % difference  [(CKD-EPI Cr 2021 − Cys 2012) / Cr 2021, %]"
PCTD_XLIM <- c(-20, 50)
# X-axis for the muscle Z-score LOESS (465 body-comp cohort).
MZ_XLAB <- "Muscle Z-score (SMI)"
MZ_XLIM <- c(-3, 2)

# ECOG distribution: % over/under-dose by ECOG group (0 / 1 / ≥2), one line per
# drug, per equation. N per ECOG group is annotated on the x-axis labels.
ecog_plot <- function(cl, eq, direction) {
  r <- reg[reg$equation == eq, ]; ttl <- paste0(r$type, " · ", r$label)
  s <- cl %>%
    dplyr::filter(equation == eq, !is.na(cls), !is.na(ecog_grp)) %>%
    dplyr::mutate(Drug = factor(drug, levels = DRUG_IDS, labels = DRUG_LABS))
  agg <- s %>% dplyr::group_by(Drug, ecog_grp) %>%
    dplyr::summarise(pct = mean(cls == direction) * 100, .groups = "drop")
  ns <- s %>% dplyr::filter(drug == DRUG_IDS[1]) %>% dplyr::count(ecog_grp)
  xl <- paste0(levels(s$ecog_grp), "\n(N=",
               ns$n[match(levels(s$ecog_grp), as.character(ns$ecog_grp))], ")")
  ggplot2::ggplot(agg, ggplot2::aes(ecog_grp, pct, colour = Drug, group = Drug)) +
    ggplot2::geom_line(linewidth = 0.9) + ggplot2::geom_point(size = 2.2) +
    ggplot2::scale_x_discrete(labels = xl) +
    ggplot2::coord_cartesian(ylim = c(0, NA)) +
    ggplot2::labs(x = "ECOG performance status", y = paste0("% ", tolower(direction), "d"),
                  title = ttl, colour = NULL) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "right",
                   plot.title = ggplot2::element_text(face = "bold", size = 11))
}

# Emit every drug × direction for one scale/version (call inside a results='asis'
# chunk). Saves each table to Word and prints it with a download link.
emit_scale <- function(cl, version, scale, scale_label, extra = FALSE, prefix = version) {
  cat("\n\n### ", scale_label, "\n\n", sep = "")
  for (d in DRUGS) {
    cat("\n\n#### ", d$lab, "\n\n", sep = "")
    for (dir in c("Overdose", "Underdose")) {
      fname <- paste0(prefix, "_", scale, "_", d$id, "_", tolower(dir))
      ft <- build_ft(cl, d$id, dir, version, extra)
      save_ft_docx(ft, fname)
      cat(flextable::flextable_to_rmd(ft))
      cat("\n\n")
    }
    cat(sprintf(
      "[📄 Overdose Word](../tables/heatmap/%1$s_%2$s_%3$s_overdose.docx) · [📄 Underdose Word](../tables/heatmap/%1$s_%2$s_%3$s_underdose.docx)\n\n",
      prefix, scale, d$id))
  }
}

# --- Most conservative GFR = per-patient MINIMUM of CKD-EPI Cr 2021,
# Cockcroft-Gault, and CKD-EPI Cr-Cys 2021 (the lowest / dose-limiting estimate).
build_mcg_long <- function(scale) {
  col_of <- function(eq) {
    r <- reg[reg$equation == eq, ]
    if (scale == "index") r$index_col[[1]] else r$nonindex_col[[1]]
  }
  mgfr <- dat[[MGFR[[scale]]]]
  mcg  <- pmin(dat[[col_of("CKDEPIcr21")]], dat[[col_of("CG")]],
               dat[[col_of("CKDEPIcrcys21")]], na.rm = TRUE)
  base <- dplyr::select(dat, sex, bmi_grp, met_grp, age_grp)
  tier <- purrr::map_dfr(TIER_DRUGS, function(dr)
    dplyr::bind_cols(base, tibble::tibble(drug = dr, cls = dose_class_tier(mcg, mgfr, dr))))
  carbo <- purrr::map_dfr(names(CARBO_BANDS), function(bn) {
    k <- CARBO_BANDS[[bn]]
    dplyr::bind_cols(base, tibble::tibble(drug = bn,
                                          cls = carbo_class(mcg, mgfr, k[["k_hi"]], k[["k_lo"]])))
  })
  dplyr::bind_rows(tier, carbo)
}
mcg_index    <- build_mcg_long("index")
mcg_nonindex <- build_mcg_long("nonindex")

# One table: rows = drugs, columns = subgroups, for the most-conservative GFR.
mcg_table <- function(cl, direction) {
  pct <- function(x) if (nrow(x)) round(mean(x$cls == direction) * 100, 1) else NA_real_
  lvl <- function(x, col, v) pct(x[!is.na(x[[col]]) & x[[col]] == v, ])
  per_drug <- function(dr) {
    e <- cl[cl$drug == dr, ]
    tibble::tibble(drug = dr, Overall = pct(e),
      Female = lvl(e, "sex", "Female"), Male = lvl(e, "sex", "Male"),
      bmi1 = lvl(e, "bmi_grp", "<20"), bmi2 = lvl(e, "bmi_grp", "20 to <25"),
      bmi3 = lvl(e, "bmi_grp", "25 to <30"), bmi4 = lvl(e, "bmi_grp", ">=30"),
      met_no = lvl(e, "met_grp", "Not metastatic"), met_yes = lvl(e, "met_grp", "Metastatic"),
      age_lt70 = lvl(e, "age_grp", "<70"), age_ge70 = lvl(e, "age_grp", ">=70"))
  }
  purrr::map_dfr(DRUG_IDS, per_drug) %>%
    dplyr::mutate(Drug = factor(drug, levels = DRUG_IDS, labels = DRUG_LABS)) %>%
    dplyr::select(Drug, dplyr::all_of(SUB_COLS))
}

# Flextable for the drug-row table (Drug label column + subgroup spanners).
make_mcg_ft <- function(tab, title, fill_fn, dark_fn) {
  ncol_total <- length(c("Drug", SUB_COLS))
  ft <- flextable::flextable(tab, col_keys = c("Drug", SUB_COLS))
  ft <- flextable::set_header_labels(ft, values = c(list(Drug = "Drug"), as.list(SUB_LABS)))
  ft <- flextable::add_header_row(ft, top = TRUE,
    values    = c("", "", "Sex", "BMI category (kg/m²)", "Metastatic disease", "Age (years)"),
    colwidths = c(1, 1, 2, 4, 2, 2))
  ft <- flextable::add_header_row(ft, top = TRUE, values = title, colwidths = ncol_total)
  for (cc in SUB_COLS) {
    v  <- tab[[cc]]
    ft <- flextable::bg(ft, j = cc, bg = fill_fn(v), part = "body")
    dark <- which(dark_fn(v))
    if (length(dark)) ft <- flextable::color(ft, i = dark, j = cc, color = "white", part = "body")
  }
  ft <- flextable::align(ft, j = SUB_COLS, align = "center", part = "all")
  ft <- flextable::align(ft, i = 2, align = "center", part = "header")
  ft <- flextable::align(ft, i = 1, align = "left",   part = "header")
  ft <- flextable::bold(ft, part = "header")
  ft <- flextable::fontsize(ft, size = 9, part = "all")
  ft <- flextable::fontsize(ft, i = 1, size = 11, part = "header")
  flextable::set_table_properties(ft, layout = "autofit", width = 1)
}

mcg_ft <- function(cl, direction, version) {
  title <- paste0(toupper(direction), " (%)")
  if (version == "cat") {
    fill_fn <- cat4_fill;                                 dark_fn <- function(x) !is.na(x) & x >= 10
  } else {
    fill_fn <- function(x) ramp_fill(x, CONT_STOPS, 30);  dark_fn <- function(x) !is.na(x) & x >= 20
  }
  make_mcg_ft(mcg_table(cl, direction), title, fill_fn, dark_fn)
}

emit_mcg <- function(cl, version, scale, scale_label) {
  cat("\n\n#### ", scale_label, "\n\n", sep = "")
  for (dir in c("Overdose", "Underdose")) {
    ft <- mcg_ft(cl, dir, version)
    save_ft_docx(ft, paste0("mcg_", version, "_", scale, "_", tolower(dir)))
    cat(flextable::flextable_to_rmd(ft)); cat("\n\n")
  }
  cat(sprintf(
    "[📄 Overdose Word](../tables/heatmap/mcg_%1$s_%2$s_overdose.docx) · [📄 Underdose Word](../tables/heatmap/mcg_%1$s_%2$s_underdose.docx)\n\n",
    version, scale))
}
```

# Overview

For every patient we compute the chemotherapy dose each **estimated-GFR (eGFR) equation** would
prescribe and compare it to the dose that **measured GFR (mGFR, ⁵¹Cr-EDTA clearance)** — the gold
standard — would prescribe. Each cell is the **percent of patients mis-dosed** in that subgroup:

- **OVERDOSE (%)** — the equation would give *more* drug than mGFR warrants.
- **UNDERDOSE (%)** — the equation would give *less* drug than mGFR warrants.

**Cell colour** — the number in each cell is the **share of patients in that subgroup whose dose
would be wrong in that direction** (over- or under-dosed vs mGFR). The same tables are shown in
**two colour versions** so you can compare:

- **Part A — categorical (4 severity bands):**
  <span style="background:#FFEDA0">&nbsp;yellow &lt; 5 %&nbsp;</span> ·
  <span style="background:#FEB24C">&nbsp;orange 5–10 %&nbsp;</span> ·
  <span style="background:#F03B20;color:white">&nbsp;red 10–20 %&nbsp;</span> ·
  <span style="background:#6A51A3;color:white">&nbsp;purple &gt; 20 %&nbsp;</span>.
- **Part B — continuous:** a shade of the value,
  <span style="background:#FFEDA0">&nbsp;yellow = 0 %&nbsp;</span> →
  <span style="background:#FEB24C">&nbsp;orange&nbsp;</span> →
  <span style="background:#B2182B;color:white">&nbsp;deep red&nbsp;</span>, saturating (worst) at
  **≥ 30 %**.

Both are severity scales (higher % = worse) applied to over- and under-dose alike; the exact
percentage is printed in every cell.

Columns give the **Overall** rate and rates within **Sex**, **BMI category**, **Metastatic
disease**, and **Age** (<70 vs ≥70) subgroups. Results are shown on both the **indexed** (mL/min/1.73 m²) and **non-indexed**
(absolute mL/min) GFR scale; mGFR is always taken on the same scale as the equation. All 1,200
patients are used for every drug (a counterfactual — the file has no administered doses).

**Dosing rules** — one row per GFR band. *Tier* is the ordered dose code used internally
(higher = more drug; 0 = avoid / forced reduce).

| Drug | GFR (mL/min) | Dose | Tier |
|---|---|---|:--:|
| **Cisplatin** | ≥ 60 | Full dose | 3 |
|  | 50 – 59 | −25 % | 2 |
|  | 40 – 49 | −50 % | 1 |
|  | < 40 | Avoid | 0 |
| **Methotrexate** (ADDIKD) | ≥ 60 | Full dose | 3 |
|  | 45 – 59.99 | −25 % | 2 |
|  | 30 – 44.99 | −50 % | 1 |
|  | < 30 | Avoid | 0 |
| **Melphalan** (200 mg/m²) | > 60 | Full, 200 mg/m² | 2 |
|  | 30 – 59 | 140 mg/m² | 1 |
|  | < 30 | Forced reduce, 140 mg/m² | 0 |
| **Carboplatin** (Calvert) | continuous (AUC) | dose = target AUC × (GFR + 25), target AUC 5; over-/under-dose set by a ±20 % or ±15 % band — **see note below** | — |

::: callout-note
## How carboplatin over- and under-dosing are defined

Carboplatin is **not** dosed in GFR tiers like the other three drugs — it is dosed to a target drug
exposure with the **Calvert formula**: **dose (mg) = target AUC × (GFR + 25)**, with target
**AUC = 5**. Because the dose is *directly proportional* to the GFR entered, plugging an eGFR
equation into the formula instead of the true (measured) GFR changes the dose by the **same
percentage** that the equation mis-estimates GFR. We flag only a *clinically meaningful* error,
using a tolerance band around the AUC-5 target:

- **Overdose** — the eGFR-based dose is larger than the dose from **measured GFR at AUC 6**. Since
  AUC 6 is 20 % above AUC 5, the equation would deliver **> 20 % too much** carboplatin.
- **Underdose** — the eGFR-based dose is smaller than the dose from **measured GFR at AUC 4** (20 %
  below AUC 5) → **> 20 % too little**.

That ±20 % band is reported as **`carboplatin_b20`**; **`carboplatin_b15`** repeats the same test
with a tighter **±15 %** band (AUC 5.75 and 4.25). Both are shown because the analysis plan asked
for 15 % and 20 % thresholds. (These definitions come from the eligibility-cutoffs document.)
:::

::: callout-note
"Metastatic disease — Unknown" from the source template is omitted because metastatic status is
complete (no missing) in this dataset. The creatinine equation shown is **CKD-EPI 2021** (the
race-free version present in the data).
:::

# Variable codebook

Every field used to build the tables. **Inputs** are columns read from
`OncoGFR1200_MGH_chemo_CJASN.xlsx` (N = 1,200); **derived** fields are computed in this analysis.

### Kidney-function inputs (eGFR / mGFR)

Each equation is stored twice — an **index** column (mL/min/1.73 m², BSA-normalised) and a
**non-index** column (absolute mL/min ≈ indexed × BSA ⁄ 1.73). Measured GFR is the reference all
equations are compared against; comparisons are made on the matching scale.

| eGFR type | Equation (label) | Index column | Non-index column |
|---|---|---|---|
| reference | Measured GFR (⁵¹Cr-EDTA) | `mGFRindex` | `mGFRnon_index` |
| eGFRcr | CKD-EPI 2021 (creatinine) | `CKDEPIcr21index` | `CKDEPIcr21non_index` |
| eGFRcr | EKFC 2021 (creatinine) | `EKFCcrindex` | `EKFCcrnon_index` |
| eGFRcys | CKD-EPI 2012 (cystatin C) | `CKDEPIcysindex` | `CKDEPIcysnon_index` |
| eGFRcys | EKFC 2023 (cystatin C) | `EKFCcysindex` | `EKFCcysnon_index` |
| eGFRcr-cys | CKD-EPI 2021 (Cr–Cys) | `CKDEPIcrcys21index` | `CKDEPIcrcys21non_index` |
| eGFRcr-cys | EKFC 2023 (Cr–Cys) | `EKFCcrcysindex` | `EKFCcrcysnon-index` |
| ClCr | Cockcroft–Gault | `CGindex` | `CGnon_index` |
| Panel | CKD-EPI 4-marker (Cr, Cys, β2M, BTP) | `CKDEPI4mark` | `CKDEPI4marknon_index` |

*Underlying labs (`Cr`, `CysC`, `B2M`, `BTP`) and anthropometrics (`weight`, `height`, `BSA`,
`age`, `gender`) were used upstream to compute these equation values; this analysis reads only the
equation columns plus `gender`, `BMI`, and `metastasis` (below).*

### Subgroup variables (table columns)

| Derived field | Source column | Levels / definition |
|---|---|---|
| `Overall` | — | all 1,200 patients |
| `sex` | `gender` (1, 2) | Male (1); Female (2) |
| `bmi_grp` | `BMI` (kg/m²) | `<20`; `20 to <25`; `25 to <30`; `≥30` |
| `met_grp` | `metastasis` (0, 1) | Not metastatic (0); Metastatic (1) |
| `age_grp` | `age` (years) | `<70`; `≥70` (binary cut-off at 70) |
| `sarc_grp` | `sarcopenia_prado` (`OncoGFR_MGH_final_0428.xlsx`, joined on `protocol`) | Non-sarcopenic (0); Sarcopenic (1). Only the **465** body-comp patients (`in_bodycomp`) — used in the 465 section. |

### Computed classification fields

| Field | Values | Meaning |
|---|---|---|
| `drug` | `cisplatin`, `methotrexate`, `melphalan`, `carboplatin_b20`, `carboplatin_b15` | which drug's dosing rule was applied. `carboplatin_b20` and `carboplatin_b15` are the **same** carboplatin analysis run with a ±20 % vs ±15 % tolerance band (see the carboplatin note under *Overview*). |
| `dose_tier` | 0–3 | ordered dose level from the rule (see *Dosing rules*); higher = more drug, 0 = avoid |
| `dose_class` (`cls`) | Overdose / Underdose / Concordant | equation's tier vs mGFR's tier (Calvert band for carboplatin) |
| `pct_over` | 0–100 | % of the subgroup classified Overdose → **OVERDOSE (%)** cells |
| `pct_under` | 0–100 | % of the subgroup classified Underdose → **UNDERDOSE (%)** cells |
| `equation` / `type` / `label` | — | equation id / eGFR family / display name (see table above) |

Cell colour: two versions of the same tables — **Part A** categorical (yellow <5 / orange 5–10 / red 10–20 / purple >20) and **Part B** continuous yellow→red (worst at ≥30 %). Higher % = worse in both; the number is printed in every cell.

# Part A — Categorical colour (4 levels)

Severity bands (same scale for over- and under-dose):
<span style="background:#FFEDA0">&nbsp;yellow &lt; 5 %&nbsp;</span> ·
<span style="background:#FEB24C">&nbsp;orange 5–10 %&nbsp;</span> ·
<span style="background:#F03B20;color:white">&nbsp;red 10–20 %&nbsp;</span> ·
<span style="background:#6A51A3;color:white">&nbsp;purple &gt; 20 %&nbsp;</span>

```{r}
#| label: partA-index
#| results: asis
emit_scale(cl_index, "cat", "index", "Indexed GFR (mL/min/1.73 m²)")
```

```{r}
#| label: partA-nonindex
#| results: asis
emit_scale(cl_nonindex, "cat", "nonindex", "Non-indexed GFR (absolute mL/min)")
```

# Part B — Continuous colour (yellow → red)

Shade ∝ value, saturating at 30 %:
<span style="background:#FFEDA0">&nbsp;yellow = 0 %&nbsp;</span> →
<span style="background:#FEB24C">&nbsp;orange&nbsp;</span> →
<span style="background:#B2182B;color:white">&nbsp;deep red ≥ 30 %&nbsp;</span> (darker = worse). The
exact % is printed in every cell.

```{r}
#| label: partB-index
#| results: asis
emit_scale(cl_index, "cont", "index", "Indexed GFR (mL/min/1.73 m²)")
```

```{r}
#| label: partB-nonindex
#| results: asis
emit_scale(cl_nonindex, "cont", "nonindex", "Non-indexed GFR (absolute mL/min)")
```


# Most conservative GFR — over/under-dose heatmap

**Most conservative GFR** = for each patient, the **lowest** of three estimates —
**CKD-EPI Cr (2021)**, **Cockcroft-Gault**, and **CKD-EPI Cr-Cys (2021)** — i.e. the dose-limiting
(safest) value. Rows = drugs, columns = subgroups; each cell is the % **over-** or **under-dosed**
vs measured GFR when this single conservative estimate drives the dose. Because it takes the lowest
GFR it gives the **lowest overdose rate of the three inputs** (over-dosing only when *all three*
still exceed mGFR), at the cost of more underdosing. Both colour versions and both GFR scales are
shown; the exact % is printed in every cell.

## Part A — Categorical colour (4 levels)

```{r}
#| label: mcg-A-index
#| results: asis
emit_mcg(mcg_index, "cat", "index", "Indexed GFR (mL/min/1.73 m²)")
```

```{r}
#| label: mcg-A-nonindex
#| results: asis
emit_mcg(mcg_nonindex, "cat", "nonindex", "Non-indexed GFR (absolute mL/min)")
```

## Part B — Continuous colour (yellow → red)

```{r}
#| label: mcg-B-index
#| results: asis
emit_mcg(mcg_index, "cont", "index", "Indexed GFR (mL/min/1.73 m²)")
```

```{r}
#| label: mcg-B-nonindex
#| results: asis
emit_mcg(mcg_nonindex, "cont", "nonindex", "Non-indexed GFR (absolute mL/min)")
```


# Body-composition subset (465) — over/under-dose heatmap

Same over/under-dose heatmaps as above but **limited to the 465 patients with body-composition
data** (matched by `protocol` to `OncoGFR_MGH_final_0428.xlsx`), with an added **Sarcopenia (Prado)**
subgroup column group (Non-sarcopenic n = 308 · Sarcopenic n = 157). Rows = equations, columns =
Overall · Sex · BMI · Metastatic · Age · **Sarcopenia**; both colour versions and both GFR scales.

## Part A — Categorical colour (4 levels)

```{r}
#| label: bc465-A-index
#| results: asis
emit_scale(cl_index_465, "cat", "index", "Indexed GFR (mL/min/1.73 m²)", extra = TRUE, prefix = "bc465_cat")
```

```{r}
#| label: bc465-A-nonindex
#| results: asis
emit_scale(cl_nonindex_465, "cat", "nonindex", "Non-indexed GFR (absolute mL/min)", extra = TRUE, prefix = "bc465_cat")
```

## Part B — Continuous colour (yellow → red)

```{r}
#| label: bc465-B-index
#| results: asis
emit_scale(cl_index_465, "cont", "index", "Indexed GFR (mL/min/1.73 m²)", extra = TRUE, prefix = "bc465_cont")
```

```{r}
#| label: bc465-B-nonindex
#| results: asis
emit_scale(cl_nonindex_465, "cont", "nonindex", "Non-indexed GFR (absolute mL/min)", extra = TRUE, prefix = "bc465_cont")
```


# LOESS — over/under-dose probability vs eGFR ratio (Cys / Cr)

For each of the **8 eGFR equations**, a **LOESS smooth** of the probability that the equation
**over-** (or **under-) doses** relative to measured GFR (mGFR), as a function of the patient's
**eGFR ratio = CKD-EPI Cys (2012) ⁄ CKD-EPI Cr (2021)** (ratio of two eGFRs, so BSA cancels and the
value is the same on either scale). **One plot per equation**, with the **5 drugs overlaid** as
colour-coded curves (shaded band = ± 95 % CI), arranged in four subsections —
**Indexed → Overdose / Underdose** and **Non-indexed → Overdose / Underdose** (**32 plots total**).
*Interpret the **CKD-EPI Cys 2012** and **CKD-EPI Cr 2021** panels with care — the x-axis ratio is
built from those two equations, so those panels are partly self-referential.*
Same OncoGFR cohort (N = 1,200). The two equation families **mirror each other**: a **low** ratio
(cystatin-eGFR below creatinine-eGFR — e.g. low muscle mass, where creatinine reads function
falsely high) makes **creatinine-based** equations overdose, whereas a **high** ratio (cystatin-eGFR
above creatinine-eGFR) makes **cystatin-based** equations overdose. So creatinine-equation overdose
curves *fall* as the ratio rises while
cystatin-equation curves *rise* — exactly what makes this eGFR ratio a useful discriminating biomarker
(underdose curves are the mirror image).

## Indexed GFR (mL/min/1.73 m²)

### Overdose

```{r}
#| label: loess-idx-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index, eq, "Overdose"))
```

### Underdose

```{r}
#| label: loess-idx-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index, eq, "Underdose"))
```

## Non-indexed GFR (absolute mL/min)

### Overdose

```{r}
#| label: loess-nix-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_nonindex, eq, "Overdose"))
```

### Underdose

```{r}
#| label: loess-nix-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_nonindex, eq, "Underdose"))
```


# LOESS (difference) — over/under-dose probability vs eGFR difference (Cys − Cr)

Identical to the LOESS section above, but the x-axis is the **eGFR difference = CKD-EPI Cys (2012)
− CKD-EPI Cr (2021)** (mL/min/1.73 m²) instead of the ratio. Negative = cystatin-eGFR below
creatinine-eGFR (creatinine reads function higher). All **8 equations**, 5 drugs overlaid,
index/non-index × over/under (**32 plots total**; interpret the CKD-EPI Cys 2012 and Cr 2021 panels
with care — the x-axis difference is built from them). As with the ratio, creatinine-equation overdose curves *fall* as the
difference rises while cystatin-equation curves *rise*.

## Indexed GFR (mL/min/1.73 m²)

### Overdose

```{r}
#| label: loessd-idx-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index, eq, "Overdose", "egfr_diff", DIFF_XLAB, DIFF_XLIM))
```

### Underdose

```{r}
#| label: loessd-idx-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index, eq, "Underdose", "egfr_diff", DIFF_XLAB, DIFF_XLIM))
```

## Non-indexed GFR (absolute mL/min)

### Overdose

```{r}
#| label: loessd-nix-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_nonindex, eq, "Overdose", "egfr_diff", DIFF_XLAB, DIFF_XLIM))
```

### Underdose

```{r}
#| label: loessd-nix-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_nonindex, eq, "Underdose", "egfr_diff", DIFF_XLAB, DIFF_XLIM))
```


# LOESS (% difference) — over/under-dose probability vs eGFR % difference

Third LOESS set: the x-axis is the **% difference = (eGFR_Cr − eGFR_Cys) ⁄ eGFR_Cr × 100**
(CKD-EPI Cr 2021 vs Cys 2012). It is a ratio of two eGFRs, so it is **scale-invariant**. **Positive**
= creatinine-eGFR above cystatin-eGFR (creatinine reads function higher). All **8 equations**, 5
drugs overlaid, index/non-index × over/under (**32 plots total**; the CKD-EPI Cys 2012 and Cr 2021
panels are partly self-referential). Because a higher % difference means creatinine reads GFR
higher, **creatinine-equation overdose curves rise** as the % difference increases while
cystatin-equation curves fall — the mirror of the ratio version.

## Indexed GFR (mL/min/1.73 m²)

### Overdose

```{r}
#| label: loessp-idx-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index, eq, "Overdose", "egfr_pctdiff", PCTD_XLAB, PCTD_XLIM))
```

### Underdose

```{r}
#| label: loessp-idx-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index, eq, "Underdose", "egfr_pctdiff", PCTD_XLAB, PCTD_XLIM))
```

## Non-indexed GFR (absolute mL/min)

### Overdose

```{r}
#| label: loessp-nix-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_nonindex, eq, "Overdose", "egfr_pctdiff", PCTD_XLAB, PCTD_XLIM))
```

### Underdose

```{r}
#| label: loessp-nix-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_nonindex, eq, "Underdose", "egfr_pctdiff", PCTD_XLAB, PCTD_XLIM))
```


# Frailty markers — muscle Z-score & ECOG

Two additional views of dosing error against markers of frailty, in the same per-equation /
5-drugs-overlaid format. Over/under-dose is always vs **measured GFR**. Indexed GFR classification.

## Muscle Z-score  (LOESS, 465 body-composition cohort)

**Y = % over- (or under-) dosed; x = muscle Z-score (SMI).** Mechanistically, lower muscle Z (less
muscle) lowers serum creatinine, so **creatinine-based** equations can read GFR falsely high →
overdose; cystatin-based equations are muscle-independent. The curves show how this plays out across
the muscle spectrum (CIs widen at the sparse extremes). N = 465; indexed GFR.

### Overdose

```{r}
#| label: mz-over
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index_465, eq, "Overdose", "muscle_z", MZ_XLAB, MZ_XLIM))
```

### Underdose

```{r}
#| label: mz-under
#| fig-width: 7
#| fig-height: 4
for (eq in LOESS_EQS) print(loess_plot(cl_index_465, eq, "Underdose", "muscle_z", MZ_XLAB, MZ_XLIM))
```

## Derived dose variables

Per-patient binary flags (**1 = yes, 0 = no**), computed **for each equation** on the indexed scale
vs measured GFR (whole 1,200 cohort):

| Variable | Definition (per equation, vs mGFR) |
|---|---|
| `cisplatin_eligible` | equation GFR ≥ 40 → cisplatin-eligible (not "avoid") |
| `carbo15_over` | Calvert dose > **15 %** above the mGFR AUC-5 dose |
| `carbo15_under` | Calvert dose > **15 %** below the mGFR AUC-5 dose |
| `carbo20_over` | Calvert dose > **20 %** above the mGFR AUC-5 dose |
| `carbo20_under` | Calvert dose > **20 %** below the mGFR AUC-5 dose |

The code that creates them (click **▸ Code** to expand):

```{r}
#| label: dose-vars
#| echo: true
dose_vars <- purrr::map_dfr(reg$equation, function(eq) {
  g <- dat[[reg$index_col[reg$equation == eq]]]   # this equation's indexed GFR
  m <- dat$mGFRindex                              # measured GFR (indexed)
  tibble::tibble(
    equation = eq, ecog_grp = dat$ecog_grp,
    cisplatin_eligible = as.integer(dose_tier(g, "cisplatin") > 0),
    carbo15_over  = as.integer(carbo_class(g, m, 5.75, 4.25) == "Overdose"),
    carbo15_under = as.integer(carbo_class(g, m, 5.75, 4.25) == "Underdose"),
    carbo20_over  = as.integer(carbo_class(g, m, 6, 4)       == "Overdose"),
    carbo20_under = as.integer(carbo_class(g, m, 6, 4)       == "Underdose"))
}) %>% dplyr::left_join(dplyr::select(reg, equation, type, label), by = "equation")
```

Overall rate (% yes) of each flag, by equation:

```{r}
#| label: dose-vars-summary
dose_vars %>%
  dplyr::transmute(Equation = paste(type, "·", label),
                   cisplatin_eligible, carbo15_over, carbo15_under, carbo20_over, carbo20_under) %>%
  dplyr::group_by(Equation) %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ round(mean(.x) * 100, 1)), .groups = "drop") %>%
  knitr::kable(col.names = c("Equation", "Cispl. eligible %", "Carbo15 over %",
                             "Carbo15 under %", "Carbo20 over %", "Carbo20 under %"))
```

## ECOG performance status  (whole 1,200 cohort)

Bars (histogram) of each dose flag by **ECOG group** (0 / 1 / ≥2; ECOG 2+3 lumped, only n = 8 are
ECOG 3), **faceted by equation**, all in one figure. N per ECOG group is under the x-axis. For
cisplatin the **ineligible** (avoid) rate is shown so it is on the same scale as the carboplatin
over/under rates.

```{r}
#| label: ecog-bars
#| fig-width: 9
#| fig-height: 12
lv <- c("Cisplatin ineligible", "Carbo15 over", "Carbo15 under", "Carbo20 over", "Carbo20 under")
pf <- dose_vars %>%
  dplyr::filter(!is.na(ecog_grp)) %>%
  dplyr::transmute(
    Equation = factor(paste(type, "·", label), levels = paste(reg$type, "·", reg$label)),
    ecog_grp,
    `Cisplatin ineligible` = 1 - cisplatin_eligible,
    `Carbo15 over` = carbo15_over, `Carbo15 under` = carbo15_under,
    `Carbo20 over` = carbo20_over, `Carbo20 under` = carbo20_under) %>%
  dplyr::group_by(Equation, ecog_grp) %>%
  dplyr::summarise(dplyr::across(dplyr::all_of(lv), ~ mean(.x) * 100), .groups = "drop") %>%
  tidyr::pivot_longer(dplyr::all_of(lv), names_to = "Outcome", values_to = "pct") %>%
  dplyr::mutate(Outcome = factor(Outcome, levels = lv))
ns <- dat %>% dplyr::filter(!is.na(ecog_grp)) %>% dplyr::count(ecog_grp)
xl <- paste0(levels(dat$ecog_grp), "\n(N=", ns$n[match(levels(dat$ecog_grp), ns$ecog_grp)], ")")
ggplot2::ggplot(pf, ggplot2::aes(ecog_grp, pct, fill = Outcome)) +
  ggplot2::geom_col(position = ggplot2::position_dodge(0.8), width = 0.75) +
  ggplot2::facet_wrap(~ Equation, ncol = 2) +
  ggplot2::scale_x_discrete(labels = xl) +
  ggplot2::labs(x = "ECOG performance status", y = "% of patients", fill = NULL,
                title = "Dose-flag rates by ECOG (whole 1,200 cohort)") +
  ggplot2::theme_minimal(base_size = 10) +
  ggplot2::theme(legend.position = "top")
```


# 5 Notes

- **Reference.** Measured GFR (⁵¹Cr-EDTA) is the gold standard; each equation is compared to mGFR on
  the matching scale (indexed vs indexed, non-indexed vs non-indexed).
- **Scale.** Calvert carboplatin dosing was derived on **absolute (non-indexed)** GFR (Section 4 is
  the clinically conventional view for carboplatin); indexed GFR (Section 3) is the conventional
  scale for CKD-staging cutoffs.
- **Direction.** Creatinine-based equations tend to **overdose** (they overestimate GFR, especially
  at low BMI / low muscle mass); cystatin-based equations tend to **underdose**.
- **Counterfactual.** No administered doses exist in this file, so every drug rule is applied to all
  1,200 patients to estimate misclassification rates, not actual prescribing errors.
