---
title: "Chemotherapy Dosing: eGFR Equations vs Measured GFR (OncoGFR 1200)"
author: "Tianqi Ouyang"
date: "2026-05-29"
execute:
  eval: true
  echo: false
  warning: false
  message: false
---

# 1 Setup {.content-hidden}

```{r}
#| label: setup-onco
# Force UTF-8 so unicode (mL/min/1.73m², ±, ≥) renders correctly.
invisible(Sys.setlocale("LC_ALL", "en_US.UTF-8"))

# Lightweight install guard so the project renders on a fresh machine.
.need <- c("readxl","dplyr","tidyr","purrr","tibble","flextable","officer")
.miss <- setdiff(.need, rownames(installed.packages()))
if (length(.miss)) install.packages(.miss, repos = "https://cloud.r-project.org")

library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(flextable)

# Output directory for downloadable Word copies of every table.
docx_dir <- "docs/tables/heatmap"
dir.create(docx_dir, recursive = TRUE, showWarnings = FALSE)

save_ft_docx <- function(ft, name) {
  flextable::save_as_docx(ft, path = file.path(docx_dir, paste0(name, ".docx")))
  invisible(ft)
}

# Traffic-light fill: green <5%, orange 5–10%, red >10% (white if NA).
tl_color <- function(x) {
  dplyr::case_when(is.na(x) ~ "#FFFFFF",
                   x <  5   ~ "#92D050",
                   x <= 10  ~ "#FFC000",
                   TRUE     ~ "#FF0000")
}
```

# 2 Data Management {.content-hidden}

```{r}
#| label: data-onco
raw <- suppressWarnings(
  readxl::read_excel("Data/OncoGFR1200_MGH_chemo_CJASN.xlsx")
)
dat <- raw %>%
  dplyr::select(-dplyr::any_of(c("...70", "...75"))) %>%
  dplyr::mutate(
    sex     = factor(gender, levels = c(1, 2), labels = c("Male", "Female")),
    bmi_grp = cut(BMI, breaks = c(-Inf, 20, 25, 30, Inf), right = FALSE,
                  labels = c("<20", "20 to <25", "25 to <30", ">=30")),
    # metastasis is complete (0/1) in this dataset — no "Unknown" level.
    met_grp = factor(metastasis, levels = c(0, 1),
                     labels = c("Not metastatic", "Metastatic"))
  )

# --- Equation registry: type + label + index/non-index columns -------------
# Reference = measured GFR. Note: the creatinine CKD-EPI column is the 2021
# (race-free) equation; the hyphen in "EKFCcrcysnon-index" is intentional.
reg <- tibble::tribble(
  ~equation,       ~type,        ~label,             ~index_col,           ~nonindex_col,
  "CKDEPIcr21",    "eGFRcr",     "CKD-EPI 2021",     "CKDEPIcr21index",    "CKDEPIcr21non_index",
  "EKFCcr",        "eGFRcr",     "EKFC 2021",        "EKFCcrindex",        "EKFCcrnon_index",
  "CKDEPIcys",     "eGFRcys",    "CKD-EPI 2012",     "CKDEPIcysindex",     "CKDEPIcysnon_index",
  "EKFCcys",       "eGFRcys",    "EKFC 2023",        "EKFCcysindex",       "EKFCcysnon_index",
  "CKDEPIcrcys21", "eGFRcr-cys", "CKD-EPI 2021",     "CKDEPIcrcys21index", "CKDEPIcrcys21non_index",
  "EKFCcrcys",     "eGFRcr-cys", "EKFC 2023",        "EKFCcrcysindex",     "EKFCcrcysnon-index",
  "CG",            "ClCr",       "Cockcroft-Gault",  "CGindex",            "CGnon_index",
  "CKDEPI4mark",   "Panel",      "CKD-EPI 4-marker", "CKDEPI4mark",        "CKDEPI4marknon_index"
)
MGFR <- c(index = "mGFRindex", nonindex = "mGFRnon_index")

# --- Dosing rules ----------------------------------------------------------
dose_tier <- function(gfr, drug) {           # higher integer = more drug; 0 = avoid
  drug <- tolower(drug)
  if (drug == "cisplatin")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr >= 60 ~ 3, gfr >= 50 ~ 2, gfr >= 40 ~ 1, TRUE ~ 0)
  else if (drug == "methotrexate")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr >= 100 ~ 2, gfr >= 40 ~ 1, TRUE ~ 0)
  else if (drug == "melphalan")
    dplyr::case_when(is.na(gfr) ~ NA_real_, gfr > 60 ~ 2, gfr >= 30 ~ 1, TRUE ~ 0)
  else stop("unknown drug: ", drug)
}
dose_class_tier <- function(gfr, mgfr, drug) {
  te <- dose_tier(gfr, drug); tr <- dose_tier(mgfr, drug)
  dplyr::case_when(is.na(te) | is.na(tr) ~ NA_character_,
                   te > tr ~ "Overdose", te < tr ~ "Underdose", TRUE ~ "Concordant")
}
# Carboplatin Calvert (AUC 5): over if 5*(eGFR+25) > k_hi*(mGFR+25); under if < k_lo.
# ±20% → k_hi=6, k_lo=4   |   ±15% → k_hi=5.75, k_lo=4.25
carbo_class <- function(gfr, mgfr, k_hi = 6, k_lo = 4) {
  pred <- 5 * (gfr + 25)
  dplyr::case_when(is.na(gfr) | is.na(mgfr) ~ NA_character_,
                   pred > k_hi * (mgfr + 25) ~ "Overdose",
                   pred < k_lo * (mgfr + 25) ~ "Underdose", TRUE ~ "Concordant")
}

TIER_DRUGS  <- c("cisplatin", "methotrexate", "melphalan")
CARBO_BANDS <- list(carboplatin_b20 = c(k_hi = 6, k_lo = 4),
                    carboplatin_b15 = c(k_hi = 5.75, k_lo = 4.25))

# Patient-level classification (carries subgroup columns) for one GFR scale.
build_class_long <- function(scale = c("index", "nonindex")) {
  scale  <- match.arg(scale)
  col_of <- function(eq) {
    r <- reg[reg$equation == eq, ]
    if (scale == "index") r$index_col[[1]] else r$nonindex_col[[1]]
  }
  mgfr <- dat[[MGFR[[scale]]]]
  base <- dplyr::select(dat, sex, bmi_grp, met_grp)
  tier <- purrr::map_dfr(reg$equation, function(eq) {
    g <- dat[[col_of(eq)]]
    purrr::map_dfr(TIER_DRUGS, function(dr)
      dplyr::bind_cols(base, tibble::tibble(equation = eq, drug = dr,
                                            cls = dose_class_tier(g, mgfr, dr))))
  })
  carbo <- purrr::map_dfr(reg$equation, function(eq) {
    g <- dat[[col_of(eq)]]
    purrr::map_dfr(names(CARBO_BANDS), function(bn) {
      k <- CARBO_BANDS[[bn]]
      dplyr::bind_cols(base, tibble::tibble(equation = eq, drug = bn,
                                            cls = carbo_class(g, mgfr, k[["k_hi"]], k[["k_lo"]])))
    })
  })
  dplyr::bind_rows(tier, carbo)
}
cl_index    <- build_class_long("index")
cl_nonindex <- build_class_long("nonindex")

# Internal (syntactic) column keys + their display labels.
SUB_COLS <- c("Overall", "Female", "Male", "bmi1", "bmi2", "bmi3", "bmi4", "met_no", "met_yes")
SUB_LABS <- c(Overall = "Overall", Female = "Female", Male = "Male",
              bmi1 = "<20", bmi2 = "20 to <25", bmi3 = "25 to <30", bmi4 = ">=30",
              met_no = "Not metastatic", met_yes = "Metastatic")

# Build the equation × subgroup % table for one drug and direction.
subgroup_table <- function(cl, drug_id, direction) {
  s <- cl %>% dplyr::filter(drug == drug_id, !is.na(cls))
  pct <- function(d) if (nrow(d)) round(mean(d$cls == direction) * 100, 1) else NA_real_
  lvl <- function(d, col, value) pct(d[!is.na(d[[col]]) & d[[col]] == value, ])
  per_eq <- function(eq) {
    e <- s %>% dplyr::filter(equation == eq)
    tibble::tibble(
      equation         = eq,
      Overall          = pct(e),
      Female           = lvl(e, "sex", "Female"),
      Male             = lvl(e, "sex", "Male"),
      bmi1    = lvl(e, "bmi_grp", "<20"),
      bmi2    = lvl(e, "bmi_grp", "20 to <25"),
      bmi3    = lvl(e, "bmi_grp", "25 to <30"),
      bmi4    = lvl(e, "bmi_grp", ">=30"),
      met_no  = lvl(e, "met_grp", "Not metastatic"),
      met_yes = lvl(e, "met_grp", "Metastatic")
    )
  }
  purrr::map_dfr(reg$equation, per_eq) %>%
    dplyr::left_join(dplyr::select(reg, equation, type, label), by = "equation") %>%
    dplyr::mutate(type = factor(type, levels = unique(reg$type))) %>%
    dplyr::arrange(type, match(equation, reg$equation)) %>%
    dplyr::select(type, label, dplyr::all_of(SUB_COLS))
}

# Render one traffic-light flextable (title e.g. "OVERDOSE (%)").
make_tl_ft <- function(tab, title) {
  ncol_total <- length(c("type", "label", SUB_COLS))   # 11
  ft <- flextable::flextable(tab, col_keys = c("type", "label", SUB_COLS))
  ft <- flextable::set_header_labels(ft,
    values = c(list(type = "eGFR type", label = "Equation"), as.list(SUB_LABS)))
  # Group-spanner header row (Sex / BMI / Metastatic disease).
  ft <- flextable::add_header_row(ft, top = TRUE,
    values    = c("", "", "", "Sex", "BMI category (kg/m²)", "Metastatic disease"),
    colwidths = c(1, 1, 1, 2, 4, 2))
  # Title row across the whole table ("OVERDOSE (%)" / "UNDERDOSE (%)").
  ft <- flextable::add_header_row(ft, top = TRUE, values = title, colwidths = ncol_total)
  for (cc in SUB_COLS) ft <- flextable::bg(ft, j = cc, bg = tl_color(tab[[cc]]), part = "body")
  ft <- flextable::merge_v(ft, j = "type", part = "body")
  ft <- flextable::valign(ft, j = "type", valign = "top", part = "body")
  ft <- flextable::align(ft, j = SUB_COLS, align = "center", part = "all")
  ft <- flextable::align(ft, i = 2, align = "center", part = "header")  # spanner row
  ft <- flextable::align(ft, i = 1, align = "left",   part = "header")  # title row
  ft <- flextable::bold(ft, part = "header")
  ft <- flextable::fontsize(ft, size = 9, part = "all")
  ft <- flextable::fontsize(ft, i = 1, size = 11, part = "header")
  # Defer column sizing to the HTML/Word renderer (layout autofit) rather than
  # autofit(), which measures text R-side and can fail under quarto's render
  # graphics context ("length = 2 in coercion to logical(1)").
  flextable::set_table_properties(ft, layout = "autofit", width = 1)
}

# Convenience: build + save + return a flextable for one (cl, drug, direction).
# Returns `ft` VISIBLY (last expression) so knitr auto-prints it in the doc.
tl <- function(cl, drug_id, direction, file) {
  ft <- make_tl_ft(subgroup_table(cl, drug_id, direction),
                   paste0(toupper(direction), " (%)"))
  save_ft_docx(ft, file)
  ft
}
```

# Overview

For every patient we compute the chemotherapy dose each **estimated-GFR (eGFR) equation** would
prescribe and compare it to the dose that **measured GFR (mGFR, ⁵¹Cr-EDTA clearance)** — the gold
standard — would prescribe. Each cell is the **percent of patients mis-dosed** in that subgroup:

- **OVERDOSE (%)** — the equation would give *more* drug than mGFR warrants.
- **UNDERDOSE (%)** — the equation would give *less* drug than mGFR warrants.

**Traffic-light colour:** <span style="background:#92D050">&nbsp;green &lt;5%&nbsp;</span> ·
<span style="background:#FFC000">&nbsp;orange 5–10%&nbsp;</span> ·
<span style="background:#FF0000;color:white">&nbsp;red &gt;10%&nbsp;</span>.

Columns give the **Overall** rate and rates within **Sex**, **BMI category**, and **Metastatic
disease** subgroups. Results are shown on both the **indexed** (mL/min/1.73 m²) and **non-indexed**
(absolute mL/min) GFR scale; mGFR is always taken on the same scale as the equation. All 1,200
patients are used for every drug (a counterfactual — the file has no administered doses).

| Drug | Dosing rule |
|---|---|
| **Cisplatin** | ≥60 full · 50–59 −25% · 40–49 −50% · <40 avoid |
| **Methotrexate** (CNS, 8 g/m²) | ≥100 full · 40–99 reduced · <40 avoid |
| **Melphalan** (200 mg/m²) | >60 full · 30–59 → 140 mg/m² · <30 forced reduce |
| **Carboplatin** (Calvert AUC 5) | over/under at AUC 6 / AUC 4 (±20%) and the ±15% margin |

::: callout-note
"Metastatic disease — Unknown" from the source template is omitted because metastatic status is
complete (no missing) in this dataset. The creatinine equation shown is **CKD-EPI 2021** (the
race-free version present in the data).
:::

# 3 Indexed GFR (mL/min/1.73 m²)

## 3.1 Carboplatin (±20%)

```{r}
#| label: idx-carbo20-over
tl(cl_index, "carboplatin_b20", "Overdose", "index_carboplatin20_over")
```

```{r}
#| label: idx-carbo20-under
tl(cl_index, "carboplatin_b20", "Underdose", "index_carboplatin20_under")
```

[📄 Overdose table](../tables/heatmap/index_carboplatin20_over.docx) ·
[📄 Underdose table](../tables/heatmap/index_carboplatin20_under.docx)

## 3.2 Carboplatin (±15%)

```{r}
#| label: idx-carbo15-over
tl(cl_index, "carboplatin_b15", "Overdose", "index_carboplatin15_over")
```

```{r}
#| label: idx-carbo15-under
tl(cl_index, "carboplatin_b15", "Underdose", "index_carboplatin15_under")
```

[📄 Overdose table](../tables/heatmap/index_carboplatin15_over.docx) ·
[📄 Underdose table](../tables/heatmap/index_carboplatin15_under.docx)

## 3.3 Cisplatin

```{r}
#| label: idx-cis-over
tl(cl_index, "cisplatin", "Overdose", "index_cisplatin_over")
```

```{r}
#| label: idx-cis-under
tl(cl_index, "cisplatin", "Underdose", "index_cisplatin_under")
```

[📄 Overdose table](../tables/heatmap/index_cisplatin_over.docx) ·
[📄 Underdose table](../tables/heatmap/index_cisplatin_under.docx)

## 3.4 Methotrexate

```{r}
#| label: idx-mtx-over
tl(cl_index, "methotrexate", "Overdose", "index_methotrexate_over")
```

```{r}
#| label: idx-mtx-under
tl(cl_index, "methotrexate", "Underdose", "index_methotrexate_under")
```

[📄 Overdose table](../tables/heatmap/index_methotrexate_over.docx) ·
[📄 Underdose table](../tables/heatmap/index_methotrexate_under.docx)

## 3.5 Melphalan

```{r}
#| label: idx-melph-over
tl(cl_index, "melphalan", "Overdose", "index_melphalan_over")
```

```{r}
#| label: idx-melph-under
tl(cl_index, "melphalan", "Underdose", "index_melphalan_under")
```

[📄 Overdose table](../tables/heatmap/index_melphalan_over.docx) ·
[📄 Underdose table](../tables/heatmap/index_melphalan_under.docx)

# 4 Non-Indexed GFR (absolute mL/min)

## 4.1 Carboplatin (±20%)

```{r}
#| label: nix-carbo20-over
tl(cl_nonindex, "carboplatin_b20", "Overdose", "nonindex_carboplatin20_over")
```

```{r}
#| label: nix-carbo20-under
tl(cl_nonindex, "carboplatin_b20", "Underdose", "nonindex_carboplatin20_under")
```

[📄 Overdose table](../tables/heatmap/nonindex_carboplatin20_over.docx) ·
[📄 Underdose table](../tables/heatmap/nonindex_carboplatin20_under.docx)

## 4.2 Carboplatin (±15%)

```{r}
#| label: nix-carbo15-over
tl(cl_nonindex, "carboplatin_b15", "Overdose", "nonindex_carboplatin15_over")
```

```{r}
#| label: nix-carbo15-under
tl(cl_nonindex, "carboplatin_b15", "Underdose", "nonindex_carboplatin15_under")
```

[📄 Overdose table](../tables/heatmap/nonindex_carboplatin15_over.docx) ·
[📄 Underdose table](../tables/heatmap/nonindex_carboplatin15_under.docx)

## 4.3 Cisplatin

```{r}
#| label: nix-cis-over
tl(cl_nonindex, "cisplatin", "Overdose", "nonindex_cisplatin_over")
```

```{r}
#| label: nix-cis-under
tl(cl_nonindex, "cisplatin", "Underdose", "nonindex_cisplatin_under")
```

[📄 Overdose table](../tables/heatmap/nonindex_cisplatin_over.docx) ·
[📄 Underdose table](../tables/heatmap/nonindex_cisplatin_under.docx)

## 4.4 Methotrexate

```{r}
#| label: nix-mtx-over
tl(cl_nonindex, "methotrexate", "Overdose", "nonindex_methotrexate_over")
```

```{r}
#| label: nix-mtx-under
tl(cl_nonindex, "methotrexate", "Underdose", "nonindex_methotrexate_under")
```

[📄 Overdose table](../tables/heatmap/nonindex_methotrexate_over.docx) ·
[📄 Underdose table](../tables/heatmap/nonindex_methotrexate_under.docx)

## 4.5 Melphalan

```{r}
#| label: nix-melph-over
tl(cl_nonindex, "melphalan", "Overdose", "nonindex_melphalan_over")
```

```{r}
#| label: nix-melph-under
tl(cl_nonindex, "melphalan", "Underdose", "nonindex_melphalan_under")
```

[📄 Overdose table](../tables/heatmap/nonindex_melphalan_over.docx) ·
[📄 Underdose table](../tables/heatmap/nonindex_melphalan_under.docx)

# 5 Notes

- **Reference.** Measured GFR (⁵¹Cr-EDTA) is the gold standard; each equation is compared to mGFR on
  the matching scale (indexed vs indexed, non-indexed vs non-indexed).
- **Scale.** Calvert carboplatin dosing was derived on **absolute (non-indexed)** GFR (Section 4 is
  the clinically conventional view for carboplatin); indexed GFR (Section 3) is the conventional
  scale for CKD-staging cutoffs.
- **Direction.** Creatinine-based equations tend to **overdose** (they overestimate GFR, especially
  at low BMI / low muscle mass); cystatin-based equations tend to **underdose**.
- **Counterfactual.** No administered doses exist in this file, so every drug rule is applied to all
  1,200 patients to estimate misclassification rates, not actual prescribing errors.
