---
title: "Grant Figures — Carboplatin Overdosing vs eGFRcr-cys / Cockcroft-Gault Discordance"
author: "Tianqi Ouyang"
execute:
  eval: true
  echo: true
  warning: false
  message: false
format:
  html:
    fig-dpi: 96
    fig-format: png
    code-fold: true
---

# Overview

A focused, grant-ready cut of the OncoGFR 1200 analysis. Two figures:

1. **Figure 1** — the probability that **carboplatin (Calvert AUC 5, ±20 % band)** is
   **overdosed** relative to measured GFR, as a smooth function of the eGFR discordance
   index. One line only (red), one panel per eGFR dosing equation.
2. **Figure 2** — how the **1,200 participants are distributed** across that same
   discordance index.

## The x-axis (Jiaxuan definition)

$$
\text{eGFR \% difference} \;=\; \frac{\text{eGFR}_{\text{cr-cys}} - \text{CrCl}_{\text{Cockcroft-Gault}}}{\text{CrCl}_{\text{Cockcroft-Gault}}} \times 100
$$

Both terms are on the **absolute mL/min scale (actual body surface area, *not* indexed to
1.73 m²)**, i.e. `CKDEPIcrcys21non_index` and `CGnon_index`. eGFR~cr-cys~ is the
**CKD-EPI 2021 creatinine–cystatin C** equation.

**Negative** = cystatin-informed eGFR reads *below* Cockcroft-Gault, i.e. Cockcroft-Gault
is flattering the kidney → the patient is at risk of receiving too much drug.
**Positive** = the reverse.

## The y-axis

Carboplatin is dosed by the Calvert formula at AUC 5. A patient is **overdosed** when the
dose computed from the estimating equation exceeds the dose that measured GFR
(⁵¹Cr-EDTA) would have given by more than **20 %**:

$$5\,(\text{eGFR} + 25) \;>\; 6\,(\text{mGFR} + 25)$$

Measured GFR is the gold standard throughout. Each panel title names the equation used to
*pick the dose*; the x-axis discordance index is the same in every panel.

# Setup and data

```{r}
#| label: grant-setup
# Force UTF-8 so unicode (mL/min/1.73m², ±, ≥) renders correctly.
invisible(Sys.setlocale("LC_ALL", "en_US.UTF-8"))

.need <- c("readxl", "dplyr", "tidyr", "purrr", "tibble", "ggplot2", "flextable", "officer", "rvg")
.miss <- setdiff(.need, rownames(installed.packages()))
if (length(.miss)) install.packages(.miss, repos = "https://cloud.r-project.org")

library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(flextable)
library(ggplot2)

# Editable PowerPoint copies of every panel, same convention as the heatmap page.
ppt_dir <- "docs/ppt"
dir.create(ppt_dir, recursive = TRUE, showWarnings = FALSE)
ppt_dir <- normalizePath(ppt_dir, mustWork = TRUE)

# Render a list of ggplots inline AND write them to an editable .pptx (one slide
# per plot) under docs/ppt/, then print a Markdown download link. Must be called
# inside a `results: asis` chunk. Editable vector via rvg::dml; raster fallback.
emit_plots <- function(plots, fname,
                       label = "⬇️ Download these panels as PowerPoint (editable)") {
  for (p in plots) print(p)
  loc <- officer::ph_location(left = 0.5, top = 1.2, width = 9, height = 5.14)  # 7:4
  doc <- officer::read_pptx()
  for (p in plots) {
    doc <- officer::add_slide(doc, layout = "Title and Content", master = "Office Theme")
    gr  <- tryCatch(rvg::dml(ggobj = p), error = function(e) NULL)
    doc <- officer::ph_with(doc, if (is.null(gr)) p else gr, location = loc)
  }
  print(doc, target = file.path(ppt_dir, paste0(fname, ".pptx")))
  cat(sprintf("\n\n[%s](../ppt/%s.pptx)\n\n", label, fname))
}
```

```{r}
#| label: grant-data
raw <- suppressWarnings(
  readxl::read_excel("Data/OncoGFR1200_MGH_chemo_CJASN.xlsx")
)

dat <- raw %>%
  dplyr::select(-dplyr::any_of(c("...70", "...75"))) %>%
  dplyr::mutate(
    # --- x-axis: Jiaxuan eGFR % difference ---------------------------------
    # (eGFRcr-cys − Cockcroft-Gault) / Cockcroft-Gault × 100.
    # BOTH on the absolute mL/min scale (actual BSA, non-indexed) — the
    # `*non_index` columns. Verified: non_index = index × BSA/1.73.
    # eGFRcr-cys = CKD-EPI 2021 creatinine–cystatin C.
    egfr_pctdiff_cg = (CKDEPIcrcys21non_index - CGnon_index) / CGnon_index * 100
  )

# --- Subgroups for the frailty / cachexia figure ----------------------------
# Fried frailty class (separate file; the key is misspelled 'protocal'); 592 of
# 1200 have it. The file ships no codebook, so the 0/1/2 ordering was checked
# against external markers: mean age rises 67.4 -> 69.4 -> 70.9 (trend p = 6e-05)
# and ECOG >=2 rises 0.0% -> 5.4% -> 27.4% (p = 2e-27). Both confirm the standard
# Fried ordering, so 0 = non-frail, 1 = pre-frail, 2 = frail.
.fr <- suppressWarnings(readxl::read_excel("Data/Frailty data to MGH.xlsx"))
dat$fried_grp <- factor(.fr$fried_classification[match(dat$protocol, .fr$protocal)],
                        levels = c(0, 1, 2),
                        labels = c("Non-frail", "Pre-frail", "Frail"))

# Nutrition markers (separate file; key = 'Protocol Number'); 689 of 1200.
# Cachexia 0 = No, 1 = Yes. ASG-PPP / PG-SGA global rating coded 1/2.
.nut <- suppressWarnings(readxl::read_excel("Data/Nutrition data.xlsx"))
names(.nut) <- trimws(names(.nut))
.nk <- match(dat$protocol, .nut$`Protocol Number`)
dat$cachexia_grp <- factor(.nut$Cachexia[.nk], levels = c(0, 1),
                           labels = c("No cachexia", "Cachexia"))
# ASG-PPP direction also established from the data, and it is the OPPOSITE of the
# intuitive "higher number = worse" reading: rating 1 is the malnourished group.
# 49.6% of rating 1 are cachectic vs 11.3% of rating 2 (Fisher p = 2e-28), with
# worse ECOG (6.2% vs 3.6% at >=2) and slightly lower BMI. Labels say so, so the
# figure cannot be misread.
dat$asgppp_grp   <- factor(.nut$`ASG-PPP`[.nk], levels = c(1, 2),
                           labels = c("Malnourished", "Well-nourished"))

# --- Equation registry: the 8 dosing equations, same as the heatmap page ----
# Reference = measured GFR. The creatinine CKD-EPI column is the 2021 (race-free)
# equation; the hyphen in "EKFCcrcysnon-index" is intentional.
reg <- tibble::tribble(
  ~equation,       ~type,        ~label,             ~index_col,           ~nonindex_col,
  "CKDEPIcr21",    "eGFRcr",     "CKD-EPI 2021",     "CKDEPIcr21index",    "CKDEPIcr21non_index",
  "EKFCcr",        "eGFRcr",     "EKFC 2021",        "EKFCcrindex",        "EKFCcrnon_index",
  "CKDEPIcys",     "eGFRcys",    "CKD-EPI 2012",     "CKDEPIcysindex",     "CKDEPIcysnon_index",
  "EKFCcys",       "eGFRcys",    "EKFC 2023",        "EKFCcysindex",       "EKFCcysnon_index",
  "CKDEPIcrcys21", "eGFRcr-cys", "CKD-EPI 2021",     "CKDEPIcrcys21index", "CKDEPIcrcys21non_index",
  "EKFCcrcys",     "eGFRcr-cys", "EKFC 2023",        "EKFCcrcysindex",     "EKFCcrcysnon-index",
  "CG",            "ClCr",       "Cockcroft-Gault",  "CGindex",            "CGnon_index",
  "CKDEPI4mark",   "Panel",      "CKD-EPI 4-marker", "CKDEPI4mark",        "CKDEPI4marknon_index"
)
MGFR <- c(index = "mGFRindex", nonindex = "mGFRnon_index")

# Carboplatin Calvert (AUC 5), ±20 % band:
#   over  if 5*(eGFR+25) > 6*(mGFR+25)
#   under if 5*(eGFR+25) < 4*(mGFR+25)
carbo_class <- function(gfr, mgfr, k_hi = 6, k_lo = 4) {
  pred <- 5 * (gfr + 25)
  dplyr::case_when(is.na(gfr) | is.na(mgfr) ~ NA_character_,
                   pred > k_hi * (mgfr + 25) ~ "Overdose",
                   pred < k_lo * (mgfr + 25) ~ "Underdose", TRUE ~ "Concordant")
}

# Patient × equation table carrying the discordance index, for one GFR scale.
build_carbo_long <- function(scale = c("nonindex", "index"), data = dat) {
  scale  <- match.arg(scale)
  mgfr   <- data[[MGFR[[scale]]]]
  col_of <- function(eq) {
    r <- reg[reg$equation == eq, ]
    if (scale == "index") r$index_col[[1]] else r$nonindex_col[[1]]
  }
  purrr::map_dfr(reg$equation, function(eq) {
    tibble::tibble(
      equation        = eq,
      egfr_pctdiff_cg = data$egfr_pctdiff_cg,
      cls             = carbo_class(data[[col_of(eq)]], mgfr)
    )
  })
}
carbo_nonindex <- build_carbo_long("nonindex")
carbo_index    <- build_carbo_long("index")

# X-axis label and window. The index runs ~ -52 % to +53 % in this cohort;
# [-50, 40] holds 98.2 % of patients and keeps the smooth off sparse tails.
PCTD_CG_XLAB <- "eGFR % difference  [(CKD-EPI Cr-Cys 2021 − Cockcroft-Gault) / Cockcroft-Gault, %]"
PCTD_CG_XLIM <- c(-50, 40)
# Ticks every 10 percentage points. The set spans the full data range (-52 to
# +53) so the unclipped distribution plots are ticked end to end; Figures 1 and 3
# are clipped to PCTD_CG_XLIM, which simply drops the breaks outside it.
PCTD_CG_BREAKS <- seq(-60, 60, 10)
# The single reference marker on every panel: a black dashed line at -10 %.
REF_LINE <- -10

# Shared x-axis scale + tick styling, so every figure on this page lines up.
pctd_x <- function() {
  list(
    ggplot2::scale_x_continuous(
      breaks = PCTD_CG_BREAKS,
      labels = paste0(PCTD_CG_BREAKS, "%"),
      minor_breaks = NULL),
    ggplot2::theme(
      axis.ticks.x        = ggplot2::element_line(colour = "black", linewidth = 0.4),
      axis.ticks.length.x = grid::unit(4, "pt"),
      panel.grid.minor.x  = ggplot2::element_blank()))
}
ref_vline <- function() {
  ggplot2::geom_vline(xintercept = REF_LINE, linetype = "dashed",
                      colour = "black", linewidth = 0.5)
}

# Carboplatin red — the same hue as the "Carboplatin (±20%)" line in the 5-drug panels.
CARBO_RED <- "#F8766D"

# One panel per equation: a single LOESS smooth of P(overdose) for carboplatin
# ±20 %, as a function of the discordance index. Binary 0/1 → % via the smooth.
grant_loess <- function(cl, eq, direction = "Overdose") {
  r   <- reg[reg$equation == eq, ]
  ttl <- paste0(r$type, " · ", r$label)
  d <- cl %>%
    dplyr::filter(equation == eq, !is.na(cls), !is.na(egfr_pctdiff_cg)) %>%
    dplyr::mutate(y = as.integer(cls == direction) * 100)

  # Fit the LOESS explicitly rather than letting geom_smooth do it, so the drawn
  # curve and its band can be clamped to [0, 100]. An unconstrained smooth of a
  # 0/1 outcome overshoots at the sparse ends (the Cockcroft-Gault panel reaches
  # ~105 % on the left), which would be indefensible in a grant figure. The fit
  # itself is untouched — only what is drawn is clipped to the possible range.
  fit  <- stats::loess(y ~ x, span = 0.9,
                       data = data.frame(x = d$egfr_pctdiff_cg, y = d$y))
  gx   <- seq(PCTD_CG_XLIM[1], PCTD_CG_XLIM[2], length.out = 200)
  pr   <- stats::predict(fit, newdata = data.frame(x = gx), se = TRUE)
  clip <- function(v) pmin(pmax(v, 0), 100)
  band <- data.frame(x   = gx,
                     fit = clip(pr$fit),
                     lo  = clip(pr$fit - 1.96 * pr$se.fit),
                     hi  = clip(pr$fit + 1.96 * pr$se.fit))

  ggplot2::ggplot(band, ggplot2::aes(x)) +
    ref_vline() +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = lo, ymax = hi),
                         fill = CARBO_RED, alpha = 0.15) +
    ggplot2::geom_line(ggplot2::aes(y = fit), colour = CARBO_RED, linewidth = 1.0) +
    pctd_x() +
    ggplot2::coord_cartesian(xlim = PCTD_CG_XLIM, ylim = c(0, 100)) +
    ggplot2::labs(x = PCTD_CG_XLAB, y = paste0("% ", tolower(direction), "d"),
                  title = ttl,
                  subtitle = "Carboplatin (±20%)") +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      plot.title    = ggplot2::element_text(face = "bold", size = 11),
      plot.subtitle = ggplot2::element_text(colour = CARBO_RED, face = "bold", size = 10))
}
```

# Figure 1 — Carboplatin (±20 %) overdosing vs the discordance index

One line per panel: **carboplatin, ±20 % band, overdose only**. Shaded band = LOESS 95 % CI.
The x-axis is ticked every **10 percentage points**; the black dashed vertical line marks
**−10 %**.

Read the panels left-to-right: as the discordance index becomes **more negative** — cystatin
pulling the estimate below Cockcroft-Gault — the risk of carboplatin overdose rises steeply
for the creatinine-only equations.

::: {.callout-note}
The **eGFRcr-cys · CKD-EPI 2021** and **ClCr · Cockcroft-Gault** panels are partly
self-referential: those two equations *define* the x-axis. Interpret them as a description
of the index itself rather than as independent evidence.
:::

## Absolute mL/min (actual BSA) — primary

This is the scale the Calvert formula requires, and it matches the x-axis definition.

```{r}
#| label: grant-carbo-nix-over
#| results: asis
#| fig-width: 7
#| fig-height: 4
emit_plots(lapply(reg$equation, function(eq) grant_loess(carbo_nonindex, eq, "Overdose")),
           "grant_carbo_pctdiffcg_nonindexed_overdose")
```

## Indexed (mL/min/1.73 m²) — secondary

Same figure with the dosing classification computed on the indexed GFR scale. The x-axis is
unchanged (still absolute mL/min, as specified).

```{r}
#| label: grant-carbo-idx-over
#| results: asis
#| fig-width: 7
#| fig-height: 4
emit_plots(lapply(reg$equation, function(eq) grant_loess(carbo_index, eq, "Overdose")),
           "grant_carbo_pctdiffcg_indexed_overdose")
```

## The same curve as numbers

The LOESS panels above in citable form: the observed carboplatin (±20 %) overdose rate
within bands of the discordance index, for the two creatinine-based equations a clinic
would actually dose from. Absolute mL/min scale, measured GFR as reference.

```{r}
#| label: grant-stratum-table
BANDS <- c(-Inf, -30, -20, -10, 0, 10, Inf)
BAND_LABS <- c("< −30 %", "−30 to −20 %", "−20 to −10 %", "−10 to 0 %",
               "0 to +10 %", "> +10 %")

stratum_tbl <- carbo_nonindex %>%
  dplyr::filter(equation %in% c("CG", "CKDEPIcr21"), !is.na(cls),
                !is.na(egfr_pctdiff_cg)) %>%
  dplyr::mutate(
    band = cut(egfr_pctdiff_cg, BANDS, labels = BAND_LABS, right = FALSE),
    eqn  = factor(equation, levels = c("CG", "CKDEPIcr21"),
                  labels = c("Cockcroft-Gault", "CKD-EPI Cr 2021"))) %>%
  dplyr::group_by(eqn, band) %>%
  dplyr::summarise(N = dplyr::n(),
                   over = sum(cls == "Overdose"), .groups = "drop") %>%
  dplyr::mutate(pct = sprintf("%.1f%%", over / N * 100)) %>%
  dplyr::select(eqn, band, N, over, pct) %>%
  tidyr::pivot_wider(names_from = eqn, values_from = c(over, pct),
                     names_glue = "{eqn}_{.value}") %>%
  # pivot_wider orders columns by VALUE (both n's, then both %'s). Regroup them
  # by EQUATION so the "Dosed by ..." header spanners sit over the right pairs.
  dplyr::select(band, N,
                `Cockcroft-Gault_over`, `Cockcroft-Gault_pct`,
                `CKD-EPI Cr 2021_over`, `CKD-EPI Cr 2021_pct`) %>%
  dplyr::rename(`eGFR % difference` = band)

stratum_tbl %>%
  flextable::flextable() %>%
  flextable::set_header_labels(
    `Cockcroft-Gault_over` = "n overdosed", `Cockcroft-Gault_pct` = "% overdosed",
    `CKD-EPI Cr 2021_over` = "n overdosed", `CKD-EPI Cr 2021_pct` = "% overdosed") %>%
  flextable::add_header_row(
    values = c("", "", "Dosed by Cockcroft-Gault", "Dosed by CKD-EPI Cr 2021"),
    colwidths = c(1, 1, 2, 2)) %>%
  flextable::align(i = 1, part = "header", align = "center") %>%
  flextable::bold(part = "header") %>%
  flextable::autofit()
```

# Figure 2 — Distribution of the discordance index across the 1,200 participants

How the cohort actually falls along the x-axis of Figure 1. Left panel: **count**
per 5-percentage-point bin. Right panel: the same as **% of the cohort**.
The black dashed line marks **−10 %**.

```{r}
#| label: grant-hist-stats
d_hist <- dat %>% dplyr::filter(!is.na(egfr_pctdiff_cg))
n_tot  <- nrow(d_hist)
q      <- quantile(d_hist$egfr_pctdiff_cg, c(0.25, 0.5, 0.75))

tibble::tibble(
  Statistic = c("N with a value", "Mean (SD)", "Median (IQR)", "Range",
                "Below 0 % (cystatin reads lower)", "At or above 0 %",
                "Below −20 %", "Below −30 %"),
  Value = c(
    format(n_tot, big.mark = ","),
    sprintf("%.1f%% (%.1f)", mean(d_hist$egfr_pctdiff_cg), sd(d_hist$egfr_pctdiff_cg)),
    sprintf("%.1f%% (%.1f to %.1f)", q[2], q[1], q[3]),
    sprintf("%.1f%% to %.1f%%", min(d_hist$egfr_pctdiff_cg), max(d_hist$egfr_pctdiff_cg)),
    sprintf("%d (%.1f%%)", sum(d_hist$egfr_pctdiff_cg <  0),
            mean(d_hist$egfr_pctdiff_cg <  0) * 100),
    sprintf("%d (%.1f%%)", sum(d_hist$egfr_pctdiff_cg >= 0),
            mean(d_hist$egfr_pctdiff_cg >= 0) * 100),
    sprintf("%d (%.1f%%)", sum(d_hist$egfr_pctdiff_cg < -20),
            mean(d_hist$egfr_pctdiff_cg < -20) * 100),
    sprintf("%d (%.1f%%)", sum(d_hist$egfr_pctdiff_cg < -30),
            mean(d_hist$egfr_pctdiff_cg < -30) * 100))
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")
```

```{r}
#| label: grant-hist
#| results: asis
#| fig-width: 7
#| fig-height: 4
BINW <- 5   # percentage points per bin

hist_base <- function(yaxis = c("count", "percent")) {
  yaxis <- match.arg(yaxis)
  ggplot2::ggplot(d_hist, ggplot2::aes(x = egfr_pctdiff_cg)) +
    ggplot2::geom_histogram(
      ggplot2::aes(y = if (yaxis == "count") ggplot2::after_stat(count)
                   else ggplot2::after_stat(count) / n_tot * 100),
      binwidth = BINW, boundary = 0,
      fill = CARBO_RED, colour = "white", linewidth = 0.3, alpha = 0.85) +
    ref_vline() +
    pctd_x() +
    ggplot2::labs(
      x = PCTD_CG_XLAB,
      y = if (yaxis == "count") "Number of participants" else "% of cohort",
      title = if (yaxis == "count")
        "Distribution of the eGFR % difference — count"
      else "Distribution of the eGFR % difference — % of cohort",
      subtitle = sprintf("OncoGFR 1200, N = %s · median %.1f%% (IQR %.1f to %.1f) · %d-point bins",
                         format(n_tot, big.mark = ","), q[2], q[1], q[3], BINW)) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      plot.title    = ggplot2::element_text(face = "bold", size = 11),
      plot.subtitle = ggplot2::element_text(size = 9, colour = "grey30"),
      panel.grid.minor = ggplot2::element_blank())
}

emit_plots(list(hist_base("count"), hist_base("percent")),
           "grant_pctdiffcg_distribution")
```

# Figure 3 — Overdose curve superimposed on the distribution

The Cockcroft-Gault overdose curve (**black, thick**, left axis) drawn over the cohort
distribution (bars, right axis), so the risk gradient and the number of patients exposed to
it are readable in one figure. Carboplatin ±20 %, absolute mL/min, measured GFR as
reference. Black dashed line at −10 %.

```{r}
#| label: grant-overlay
#| results: asis
#| fig-width: 7.5
#| fig-height: 4.4

# Bin counts set the right-hand axis; round up to a clean tick interval.
.hc     <- graphics::hist(d_hist$egfr_pctdiff_cg,
                          breaks = seq(-60, 60, BINW), plot = FALSE)$counts
CNT_MAX <- ceiling(max(.hc) / 25) * 25

# Cockcroft-Gault overdose LOESS, clipped to [0, 100] exactly as in Figure 1.
.d_cg <- carbo_nonindex %>%
  dplyr::filter(equation == "CG", !is.na(cls), !is.na(egfr_pctdiff_cg)) %>%
  dplyr::mutate(y = as.integer(cls == "Overdose") * 100)
.fit  <- stats::loess(y ~ x, span = 0.9,
                      data = data.frame(x = .d_cg$egfr_pctdiff_cg, y = .d_cg$y))
.gx   <- seq(PCTD_CG_XLIM[1], PCTD_CG_XLIM[2], length.out = 200)
.pr   <- stats::predict(.fit, newdata = data.frame(x = .gx), se = TRUE)
.clip <- function(v) pmin(pmax(v, 0), 100)
ov <- data.frame(x  = .gx,
                 fit = .clip(.pr$fit),
                 lo  = .clip(.pr$fit - 1.96 * .pr$se.fit),
                 hi  = .clip(.pr$fit + 1.96 * .pr$se.fit))

# Left axis = % overdosed (0-100). The bars are rescaled onto that range and the
# right axis inverts the same transform, so both series share one panel.
overlay_plot <- function(yaxis = c("count", "percent")) {
  yaxis <- match.arg(yaxis)
  # Bar heights are the same either way — every bin is drawn as a fraction of
  # CNT_MAX on the 0-100 left axis. Only the right-hand axis relabels that same
  # height, either as a raw count or as a share of the cohort.
  sec_tr   <- if (yaxis == "count") ~ . / 100 * CNT_MAX
              else ~ . / 100 * CNT_MAX / n_tot * 100
  sec_name <- if (yaxis == "count") "Number of participants" else "% of cohort"
  sec_brk  <- if (yaxis == "count") seq(0, CNT_MAX, 25) else ggplot2::waiver()

  ggplot2::ggplot() +
    ggplot2::geom_histogram(
      data = d_hist,
      ggplot2::aes(x = egfr_pctdiff_cg,
                   y = ggplot2::after_stat(count) / CNT_MAX * 100),
      binwidth = BINW, boundary = 0,
      fill = CARBO_RED, colour = "white", linewidth = 0.3, alpha = 0.55) +
    ref_vline() +
    ggplot2::geom_ribbon(data = ov, ggplot2::aes(x = x, ymin = lo, ymax = hi),
                         fill = "black", alpha = 0.12) +
    ggplot2::geom_line(data = ov, ggplot2::aes(x = x, y = fit),
                       colour = "black", linewidth = 1.7) +
    pctd_x() +
    ggplot2::scale_y_continuous(
      name   = "% overdosed  (carboplatin ±20%, dosed by Cockcroft-Gault)",
      breaks = seq(0, 100, 25),
      sec.axis = ggplot2::sec_axis(sec_tr, name = sec_name, breaks = sec_brk)) +
    ggplot2::coord_cartesian(xlim = PCTD_CG_XLIM, ylim = c(0, 100)) +
    ggplot2::labs(
      x = PCTD_CG_XLAB,
      title = "Carboplatin overdose risk and cohort distribution",
      subtitle = sprintf("Black line = %% overdosed (left axis) · bars = distribution (right axis) · N = %s",
                         format(n_tot, big.mark = ","))) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      plot.title       = ggplot2::element_text(face = "bold", size = 11),
      plot.subtitle    = ggplot2::element_text(size = 9, colour = "grey30"),
      axis.title.y.left  = ggplot2::element_text(colour = "black"),
      axis.title.y.right = ggplot2::element_text(colour = CARBO_RED),
      axis.text.y.right  = ggplot2::element_text(colour = CARBO_RED),
      panel.grid.minor   = ggplot2::element_blank())
}

emit_plots(list(overlay_plot("count"), overlay_plot("percent")),
           "grant_overlay_cg_overdose_distribution")
```

# Figure 4 — Overdose risk at 10 %, 15 % and 20 %, by frailty and cachexia

Dose set from **eGFR~cr-cys~ (CKD-EPI 2021)**, reference = measured GFR, absolute mL/min.
A patient is overdosed at threshold *t* when $5(\text{eGFR}+25) > 5(1+t)(\text{mGFR}+25)$.
Bars are exact binomial 95 % CIs. The dashed line is the whole-cohort rate at each threshold.

::: {.callout-warning}
## What this figure actually shows

The drafted sentence — *"Among subset of patients who underwent frailty and cachexia
assessments, the risk of overdosing increased"* — is **not supported** by these data, on
either GFR scale. Read either way:

- **Assessed subset vs whole cohort.** The frailty-assessed subset is *lower* than the full
  cohort at every threshold (≥20 %: 9.0 % vs 11.8 %). The cachexia-assessed subset is
  indistinguishable from it (12.8 % vs 11.8 %).
- **Within the subsets.** Fried frailty shows only a weak, non-significant gradient
  (≥15 %: 12.1 % → 17.8 % → 19.5 % across non-frail / pre-frail / frail, trend *p* = 0.08;
  the ≥20 % row is flat). The frailty ordering itself is sound — it was verified against age
  and ECOG, both of which rise monotonically across the three groups. Cachexia shows nothing
  at all (≥20 %: 12.4 % with vs 13.0 % without, *p* = 0.90).

- **ASG-PPP (PG-SGA)** is the only marker with a visible separation — ≥20 %: 15.6 % vs
  10.6 % — but it runs **backwards**. Neither source file carries a codebook, so the coding
  was established from the data: rating **1 is the malnourished group** (49.6 % of them
  cachectic vs 11.3 % of rating 2, Fisher *p* = 2 × 10⁻²⁸; also worse ECOG, lower BMI).
  The group being overdosed more often is therefore the **well-nourished** one — the
  opposite of the hypothesis. It is also not significant (*p* = 0.065). All confidence
  intervals overlap.

The headline threshold result is solid and worth keeping: dosing from eGFR~cr-cys~ instead
of measured GFR still overdoses **31.9 %** of the cohort by ≥10 %, **19.8 %** by ≥15 % and
**11.8 %** by ≥20 %. The frailty/cachexia amplification claim needs rewording before it
goes in the grant.
:::

```{r}
#| label: grant-thresholds
#| results: asis
#| fig-width: 8
#| fig-height: 4.6
THRESHOLDS <- c(0.10, 0.15, 0.20)
THR_LABS   <- c("≥10% overdose", "≥15% overdose", "≥20% overdose")
THR_COLS   <- c("#FDD0A2", "#FD8D3C", "#D94801")

# Overdose indicator for an arbitrary threshold (Calvert AUC 5).
carbo_over_thr <- function(gfr, mgfr, thr)
  ifelse(is.na(gfr) | is.na(mgfr), NA_integer_,
         as.integer(5 * (gfr + 25) > 5 * (1 + thr) * (mgfr + 25)))

g_crcys <- dat$CKDEPIcrcys21non_index
m_ref   <- dat$mGFRnon_index

.domains <- list(
  list(domain = "Whole cohort",
       grp = factor(rep("Overall", nrow(dat)), levels = "Overall")),
  list(domain = "Fried frailty",     grp = dat$fried_grp),
  list(domain = "Cachexia",          grp = dat$cachexia_grp),
  list(domain = "ASG-PPP (PG-SGA)",  grp = dat$asgppp_grp))

sg <- purrr::map_dfr(.domains, function(d)
  purrr::map_dfr(seq_along(THRESHOLDS), function(i) {
    o  <- carbo_over_thr(g_crcys, m_ref, THRESHOLDS[i])
    ok <- !is.na(o) & !is.na(d$grp)
    # grp_i carries the factor's own level order; bind_rows across domains would
    # otherwise collapse the factors to character and sort them alphabetically
    # ("Frail, Non-frail, Pre-frail" instead of the clinical ordering).
    tibble::tibble(domain = d$domain, grp = as.character(d$grp[ok]),
                   grp_i = as.integer(d$grp[ok]), o = o[ok],
                   thr = THR_LABS[i]) %>%
      dplyr::group_by(domain, grp, grp_i, thr) %>%
      dplyr::summarise(N = dplyr::n(), k = sum(o), .groups = "drop")
  }))

# Exact binomial CI per cell.
.ci <- mapply(function(k, n) stats::binom.test(k, n)$conf.int, sg$k, sg$N)
sg <- sg %>% dplyr::mutate(
  pct = k / N * 100, lo = .ci[1, ] * 100, hi = .ci[2, ] * 100,
  thr = factor(thr, levels = THR_LABS),
  domain = factor(domain, levels = c("Whole cohort", "Fried frailty",
                                     "Cachexia", "ASG-PPP (PG-SGA)")),
  xlab = paste0(grp, "\n(N=", N, ")"))

# Lock the bar order to domain, then the group's own factor level.
.lev <- sg %>% dplyr::distinct(domain, grp_i, xlab) %>%
  dplyr::arrange(domain, grp_i) %>% dplyr::pull(xlab)
sg$xlab <- factor(sg$xlab, levels = .lev)

# Whole-cohort reference rate per threshold, drawn as a dashed rule.
ref <- sg %>% dplyr::filter(domain == "Whole cohort") %>%
  dplyr::select(thr, ref = pct)

p_thr <- ggplot2::ggplot(sg, ggplot2::aes(xlab, pct, fill = thr)) +
  ggplot2::geom_hline(data = ref, ggplot2::aes(yintercept = ref),
                      linetype = "dashed", colour = "grey45", linewidth = 0.4) +
  ggplot2::geom_col(position = ggplot2::position_dodge(width = 0.8), width = 0.72) +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = lo, ymax = hi),
                         position = ggplot2::position_dodge(width = 0.8),
                         width = 0.18, linewidth = 0.4, colour = "grey25") +
  ggplot2::facet_grid(thr ~ domain, scales = "free_x", space = "free_x") +
  ggplot2::scale_fill_manual(values = stats::setNames(THR_COLS, THR_LABS)) +
  ggplot2::labs(
    x = NULL, y = "% overdosed", fill = NULL,
    title = "Carboplatin overdose risk when dosing by eGFRcr-cys, by frailty and cachexia",
    subtitle = "CKD-EPI 2021 cr-cys vs measured GFR · absolute mL/min · exact binomial 95% CI\nDashed rule = whole-cohort rate at that threshold") +
  ggplot2::theme_minimal(base_size = 10) +
  ggplot2::theme(
    legend.position = "none",
    plot.title      = ggplot2::element_text(face = "bold", size = 11),
    plot.subtitle   = ggplot2::element_text(size = 8.5, colour = "grey30"),
    strip.text      = ggplot2::element_text(face = "bold", size = 9),
    panel.grid.major.x = ggplot2::element_blank(),
    axis.text.x     = ggplot2::element_text(size = 8))

emit_plots(list(p_thr), "grant_crcys_overdose_thresholds_frailty_cachexia")
```

```{r}
#| label: grant-thresholds-table
# One block per assessment domain, each closed by its own significance row, so a
# reader compares groups down a block rather than across a flat list.
#   3 ordered groups -> Cochran-Armitage trend test (prop.trend.test)
#   2 groups         -> Fisher exact test
pval_row <- function(d) {
  rows <- sg %>% dplyr::filter(domain == d)
  ps <- vapply(THR_LABS, function(t) {
    r <- rows %>% dplyr::filter(thr == t) %>% dplyr::arrange(grp_i)
    if (nrow(r) == 3) stats::prop.trend.test(r$k, r$N)$p.value
    else stats::fisher.test(matrix(c(r$k[1], r$N[1] - r$k[1],
                                     r$k[2], r$N[2] - r$k[2]), nrow = 2))$p.value
  }, numeric(1))
  tibble::tibble(
    Domain = d,
    Group  = if (nrow(rows) / length(THR_LABS) == 3) "p (trend)" else "p (difference)",
    N = "", `≥10%` = sprintf("%.2f", ps[1]),
    `≥15%` = sprintf("%.2f", ps[2]), `≥20%` = sprintf("%.2f", ps[3]))
}

cells <- sg %>%
  dplyr::arrange(domain, grp_i, thr) %>%
  dplyr::transmute(Domain = domain,
                   Group = ifelse(grp == "Overall", "Overall (reference)", grp),
                   N = format(N, big.mark = ","), thr,
                   v = sprintf("%.1f (%.1f–%.1f)", pct, lo, hi)) %>%
  tidyr::pivot_wider(names_from = thr, values_from = v) %>%
  dplyr::rename(`≥10%` = !!THR_LABS[1], `≥15%` = !!THR_LABS[2], `≥20%` = !!THR_LABS[3])

# Interleave each domain's groups with its significance row.
DOMS <- c("Fried frailty", "Cachexia", "ASG-PPP (PG-SGA)")
tbl <- dplyr::bind_rows(
  cells %>% dplyr::filter(Domain == "Whole cohort"),
  purrr::map_dfr(DOMS, ~ dplyr::bind_rows(
    cells %>% dplyr::filter(Domain == .x), pval_row(.x)))) %>%
  dplyr::mutate(Domain = as.character(Domain))

.p_rows <- which(grepl("^p \\(", tbl$Group))
.ref_row <- which(tbl$Group == "Overall (reference)")
.starts  <- which(!duplicated(tbl$Domain))

flextable::flextable(tbl) %>%
  flextable::add_header_row(
    values = c("", "", "", "% of group overdosed (95% CI)"),
    colwidths = c(1, 1, 1, 3)) %>%
  flextable::merge_v(j = "Domain") %>%
  flextable::valign(j = "Domain", valign = "top") %>%
  flextable::bold(part = "header") %>%
  flextable::align(i = 1, part = "header", align = "center") %>%
  flextable::align(j = c("N", "≥10%", "≥15%", "≥20%"), align = "right") %>%
  flextable::bold(i = .ref_row) %>%
  flextable::bg(i = .ref_row, bg = "#F2F2F2") %>%
  flextable::italic(i = .p_rows) %>%
  flextable::color(i = .p_rows, color = "grey35") %>%
  flextable::hline(i = setdiff(.starts - 1, 0)) %>%
  flextable::width(j = "Domain", width = 1.25) %>%
  flextable::width(j = "Group", width = 1.5) %>%
  flextable::width(j = "N", width = 0.5) %>%
  flextable::width(j = c("≥10%", "≥15%", "≥20%"), width = 1.35) %>%
  flextable::fontsize(size = 9, part = "all") %>%
  flextable::padding(padding.top = 3, padding.bottom = 3, part = "all") %>%
  flextable::add_footer_lines(
    "Read down each block: groups within a domain, closed by a test of whether they differ. Trend test across the three ordered frailty groups; Fisher exact for two-group comparisons. Reference row shaded. ASG-PPP rating 1 is the malnourished group (see the callout above) — the higher overdose rate sits with the well-nourished.") %>%
  flextable::fontsize(size = 8, part = "footer") %>%
  flextable::color(part = "footer", color = "grey35")
```

# Notes

- **Cohort**: OncoGFR 1200 — 1,200 patients with a measured GFR (⁵¹Cr-EDTA plasma
  clearance) as the reference standard. The discordance index is available for all 1,200.
- **Equation choices**: eGFR~cr-cys~ on the x-axis is CKD-EPI 2021 creatinine–cystatin C.
  Swap `CKDEPIcrcys21non_index` for `EKFCcrcysnon-index` in the `grant-data` chunk to use
  EKFC 2023 instead.
- **Scale**: the x-axis uses absolute mL/min (actual BSA) for both terms, as specified.
  Because the index is a *ratio* of two GFRs on the same scale, the BSA/1.73 factor cancels
  algebraically, so an indexed version of the x-axis would carry the same meaning; in the
  stored columns the two agree to within 0.52 percentage points (rounding only, r = 0.99999).
- **LOESS**: `span = 0.9`, binary over/not-over outcome smoothed on the 0/100 scale; the
  shaded band is the smoother's 95 % CI (fit ± 1.96 SE), not a binomial CI. The curve and
  band are **clipped to [0, 100]** for display — an unconstrained LOESS of a 0/1 outcome
  overshoots where data are sparse. The underlying fit is not modified, so a curve that sits
  flat against 0 % or 100 % at the extreme ends is at the edge of the data, not a plateau.
- **Variable codebook**: every input column, derived variable, function and constant used
  on this page — plus the concepts behind them (Calvert, indexed vs absolute GFR, LOESS
  clipping, exact binomial CIs) — is documented in the
  [codebook appendix on the Summary page](../index.qmd#appendix-variable-codebook-logic-and-concepts).
- Full equation-by-drug heatmaps and the other LOESS sets live on the
  [Dosing Heatmaps](dosing_heatmap.qmd) page.
